### 2019-12-17 interim/interim.tex
Add LaTeX build directory to .gitignore

\documentclass[12pt,twoside]{report}

\newcommand{\reporttitle}{Type-safe Webservices Generation: Interim Report}
\newcommand{\reportauthor}{Anson Miu}
\newcommand{\reporttype}{MEng Individual Project}
\newcommand{\supervisor}{Prof. Nobuko Yoshida}
\newcommand{\sndmarker}{TBC}
\newcommand{\cid}{your college-id number}

% include files that load packages and define macros
\input{includes} % various packages needed for maths etc.
\input{notation} % short-hand notation and macros

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
% front page
\input{titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\doublespacing
\tableofcontents
\singlespacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Main document

\chapter{Introduction}

\chapter{Background}

\chapter{Project Plan}
\section{Theory}
\section{Implementation}
\section{Timetable}

\chapter{Evaluation Plan}
\section{Theory}
\section{Implementation}

\clearpage

\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: