### 2019-12-20 interim/sections/background/codegen.tex
Start setting up interim scaffolding

\section{Code Generation}

\subsection{Overview}
\begin{itemize}
\item Motivate the practical implementation of session types in programming
\item What needs to be verified - I/O behaviour (introduce the term behaviour typing) and linear channel usage
\end{itemize}

\subsection{Static Session Typing}
\begin{itemize}
\item PureScript implementation that addresses the 2 criteria statically
\item \begin{itemize}
\item Relevance for web development
\item Static behavioural typing
\item Static channel linearity by construction
\item Pros - type dependencies (expressing dependent types)
\item Cons
\end{itemize}
\item Other initiatives -- creating language with first-class channel primitives (e.g. SILL)
\end{itemize}

PureScript - \cite{PS2019}

\subsection{Runtime Monitors}

\subsection{Hybrid Session Verification}
\begin{itemize}
\item How the aforementioned 2 approaches motivated hybrid session verification
\item Workflow using Scribble toolchain 
\item Static behavioural typing, dynamic channel usage runtime checks
\item Pros - Abstract I/O interfaces
\item Pros - Input futures for `complex communication patterns' motivated by SMTP example
\item Cons - ``practical compromise'' as outlined in paper
\end{itemize}

Java - \cite{HY2016}

\subsection{Comparison}
\begin{itemize}
\item Comparison table between the 3 approaches for how they deal with behavioural typing 
\end{itemize}