### 2019-12-20 interim/sections/background/session_types.tex

\section{Session Types}

\subsection{Overview}
\begin{itemize}
\item Motivate different ways of modelling concurrency (shared memory, message passing), narrow focus to message passing
\item Outline relevance of process algebra and session types as the relevant typing discipline
\item Outline existing implementations of session types in practice
\end{itemize}

\subsection{Process Calculus}
\begin{itemize}
\item Introduce asynchronous pi calculus
\item Motivate how this can model interaction between concurrent processes through reduction rule in operational semantics
\end{itemize}

\subsection{Binary Session Types}
\begin{itemize}
\item Introduce synchronous pi calculus with branching and selection
\item Motivate with travel agency example
\item Introduce binary session types, notion of duality and subtyping for well-typed binary proesses
\item Guarantees for well-typed sessions?
\end{itemize}

\subsection{Multiparty session types}
\begin{itemize}
\item Motivate the extension from BST to MPST (example of battleship game, existing distributed protocols that involve multiple participants?)
\item Introduce global type, projection
\item How does the notion of compatibility (which used to be duality) adapt to MPST?
\end{itemize}