### 2019-12-26 interim/interim.tex
Interim write-up: session type pi calculus

\documentclass[12pt,twoside]{report}

\newcommand{\reporttitle}{Type-safe Webservices Generation: Interim Report}
\newcommand{\reportauthor}{Anson Miu}
\newcommand{\reporttype}{MEng Individual Project}
\newcommand{\supervisor}{Prof. Nobuko Yoshida}
\newcommand{\sndmarker}{TBC}
\newcommand{\cid}{your college-id number}

% include files that load packages and define macros
\input{includes} % various packages needed for maths etc.
\input{notation} % short-hand notation and macros

\usepackage[nottoc]{tocbibind}
\setcitestyle{numbers}

\input{custom_includes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
% front page
\input{titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\doublespacing
\tableofcontents
%\singlespacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Main document

\chapter{Introduction}

% Rise of distributed programs -> main challenges

% How to ensure correctness in general: type system, data types

% Type discipline for concurrent programs - behavioural types (--> session types)

% Mainstream example of distirbuted programs: web services (microservice architecture), interactive web application;

% Objective:

\chapter{Background}

\input{sections/background/session_types.tex}
\clearpage

\input{sections/background/scribble.tex}
\clearpage

\input{sections/background/codegen.tex}
\clearpage

\input{sections/background/typescript.tex}

\chapter{Project Plan}
\section{Theory}
\section{Implementation}
\section{Timetable}

\chapter{Evaluation Plan}
\section{Theory}
\section{Implementation}

\clearpage

\bibliographystyle{acm}
\bibliography{interim}

\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: