### 2020-01-23 interim/interim.tex
Interim: update title page and extract sections as separate .tex files

\documentclass[12pt,twoside]{report}

\newcommand{\reporttitle}{Type-safe Webservices Generation: Interim Report}
\newcommand{\reportauthor}{Anson Miu}
\newcommand{\reporttype}{MEng Individual Project}
\newcommand{\supervisor}{Prof. Nobuko Yoshida}
\newcommand{\sndmarker}{Dr. Iain Phillips}
\newcommand{\cid}{your college-id number}

% include files that load packages and define macros
\input{includes} % various packages needed for maths etc.
\input{notation} % short-hand notation and macros

\usepackage[nottoc]{tocbibind}
\setcitestyle{numbers}

\input{custom_includes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
% front page
\input{titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\doublespacing
\tableofcontents
%\singlespacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Main document

\input{sections/introduction.tex}

\chapter{Background}

\input{sections/background/session_types.tex}

\input{sections/background/scribble.tex}

\input{sections/background/codegen.tex}

%\input{sections/background/typescript.tex}

\input{sections/projectplan.tex}

\input{sections/evaluationplan.tex}

\clearpage

\bibliographystyle{acm}
\bibliography{interim}

\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: