### 2020-01-24 PLACES_2020/conclusion.tex

\section{Conclusion and Future Works}
We have presented a MPST-based framework for developing full-stack interactive
TypeScript applications with WebSocket communications that conform to a
protocol, statically providing linear channel usage guarantees and affine
channel usage guarantees for back-end and front-end targets respectively.

Future works include incorporating \textit{explicit connection actions}
introduced in \cite{ExplicitConnections} in our API generation to better model
real-world communication protocols that may feature in interactive web
applications.
Whilst our approach supports multiparty sessions, the nature of
WebSockets require some server-based role in the communication protocol and
clients to interact via the server.
Extending support to WebRTC \fz{cite it or we don't know what it is} would
cater for peer-to-peer communication between browsers, which further opens up
possibilities for communication protocols supported by our approach.

% Future work
% - generalise to encode protocols as typestates (e.g. Mungo example)
% - WebRTC for full MPST (since current limitation is w.r.t. MPST protocols with 1-server many-client roles