### 2020-03-08 PLACES_2020/casestudy.tex
PLACES: update case study code listing

\section{Case Study}
\label{section:example}

\begin{figure}
\begin{lstlisting}[language=JavaScript, tabsize=4]
const handleP1Move: S13 = (move: Point) => {
	board.P1(move);			// User logic
	if (board.won()) {
		return [Labels.Lose, move, [Labels.Win, move]]; (*@\label{linelose}@*)
	} else if (board.draw()) {
		return [Labels.Draw, move, [Labels.Draw, move]]; (*@\label{linedraw}@*)
	} else {
		return [Labels.Update, move, [Labels.Update, move, handleP2Move]]; 	(*@\label{lineupdate}@*)
	}
}

// Instantiate session - `handleP2Move` defined similarly as S18
new NoughtsAndCrosses.Svr(webSocketServer, handleP1Move);
\end{lstlisting}
\captionof{lstlisting}{Session runtime instantiation for \textit{Noughts and Crosses} \texttt{Svr}.}
\label{lst:svrprotocol}
\end{figure}

We apply our framework to implement a web-based implementation of the
\textit{Noughts and Crosses} running example in TypeScript;
the interested reader can find the full implementation in
\cite{NoughtsAndCrosses}.
In addition to MPST-safety, we show that our library design welcomes idiomatic
JavaScript practices in the user implementation and is interoperable with
common front- and back-end frameworks.

\paragraph{Game Server}
We set up the WebSocket server as an Express.js \cite{ExpressJS}
application on top of a Node.js \cite{NodeJS} runtime.
We define our own game logic in a \texttt{Board} class to keep track of the
game state and expose methods to query the result.
This custom logic is integrated into our \texttt{handleP1Move} and
\texttt{handleP2Move} handlers (\cref{lst:svrprotocol}), so the session runtime
can handle \texttt{Pos(Point)} messages from players and transition to the 
permitted successor states (\cref{lst:game}) according
to the injected game logic: if \texttt{P1} played a winning move (\cref{linelose}), 
\texttt{Svr} sends a \texttt{Lose} message to \texttt{P2} with the winning 
move, and also sends a \texttt{Win} message to \texttt{P1}; 
if \texttt{P1}'s move resulted 
in a draw (\cref{linedraw}), \texttt{Svr} sends \texttt{Draw} messages to 
both \texttt{P2} and \texttt{P1}; otherwise, the game continues (\cref{lineupdate}), 
so \texttt{Svr} updates both \texttt{P2} and \texttt{P1} with the latest 
move and proceeds to handle \texttt{P2}'s turn.

Note that, by TypeScript's structural typing \cite{TypeScriptSpec}, replacing
\texttt{handleP2Move} on \cref{lineupdate} with a recursion of \texttt{handleP1Move} would be
type-correct -- this allows for better code reuse as opposed to defining
additional abstractions to work around the limitations of nominal typing in
\cite{Hybrid2016}.
There is also full type erasure when transpiling to JavaScript to run the
server code, so state space explosion is not a runtime consideration.

\paragraph{Game Clients}
We implement the game client for \texttt{P1} and
\texttt{P2} by extending from the generated abstract React (EFSM state)
components and registering those to the session runtime component.

For the sake of code reuse for demonstration purposes, \cite{NoughtsAndCrosses}
uses \textit{higher-order components} (HOC) to build the correct state
implementations depending on which player the user chooses to be.
We leverage the \textit{Redux} \cite{Redux} state management library to keep track of game
state, thus showing the flexibility of our library design in being
interoperable with other libraries and idiomatic JavaScript practices.
% REVIEW: tight-coupling?
Our approach encourages the separation of concerns between the 
communication logic and program logic -- the generated session runtime
keeps track of the state of the EFSM to ensure protocol conformance by 
construction, whilst \textit{Redux} solely manages our game state.