### 2020-03-09 PLACES_2020/casestudy.tex
More proofreading

\section{Case Study}
\label{section:example}

\begin{figure}
\begin{lstlisting}[language=JavaScript, tabsize=4]
const handleP1Move: S13 = (move: Point) => {
	board.P1(move);			// User logic
	if (board.won()) {
		return [Labels.Lose, move, [Labels.Win, move]]; (*@\label{linelose}@*)
	} else if (board.draw()) {
		return [Labels.Draw, move, [Labels.Draw, move]]; (*@\label{linedraw}@*)
	} else {
		return [Labels.Update, move, [Labels.Update, move, handleP2Move]]; 	(*@\label{lineupdate}@*)
	}
}

// Instantiate session - `handleP2Move` defined similarly as S18
new NoughtsAndCrosses.Svr(webSocketServer, handleP1Move);
\end{lstlisting}
\captionof{lstlisting}{Session runtime instantiation for \textit{Noughts and Crosses} \texttt{Svr}.}
\label{lst:svrprotocol}
\end{figure}

We apply our framework to implement a web-based implementation of the
\textit{Noughts and Crosses} running example in TypeScript;
the interested reader can find the full implementation in
\cite{NoughtsAndCrosses}.
In addition to MPST-safety, we show that our library design welcomes idiomatic
JavaScript practices in the user implementation and is interoperable with
common front- and back-end frameworks.

\paragraph{Game Server}
We set up the WebSocket server as an Express.js \cite{ExpressJS}
application on top of a Node.js \cite{NodeJS} runtime.
We define our own game logic in a \texttt{Board} class to keep track of the
game state and expose methods to query the result.
This custom logic is integrated into our \texttt{handleP1Move} and
\texttt{handleP2Move} handlers (\cref{lst:svrprotocol}), so the session runtime
can handle \texttt{Pos(Point)} messages from players and transition to the
permitted successor states (\cref{lst:game}) according
to the injected game logic: if \texttt{P1} played a winning move (\cref{linelose}),
\texttt{Svr} sends a \texttt{Lose} message to \texttt{P2} with the winning
move, and also sends a \texttt{Win} message to \texttt{P1};
if \texttt{P1}'s move resulted
in a draw (\cref{linedraw}), \texttt{Svr} sends \texttt{Draw} messages to
both \texttt{P2} and \texttt{P1}; otherwise, the game continues (\cref{lineupdate}),
so \texttt{Svr} updates both \texttt{P2} and \texttt{P1} with the latest
move and proceeds to handle \texttt{P2}'s turn.

Note that, by TypeScript's structural typing
\cite{TypeScriptSpec}, replacing
\texttt{handleP2Move} on \cref{lineupdate} with a recursive
occurrence of \texttt{handleP1Move} would be
type-correct --- this allows for better code reuse as opposed to
defining
additional abstractions to work around the limitations of nominal
typing in
\cite{Hybrid2016}.
There is also full type erasure when transpiling to JavaScript to
run the
server code, so the types defined in TypeScript will not appear in the
JavaScript after typechecking.
This means state space explosion is not a runtime consideration.

\paragraph{Game Clients}
We implement the game client for \texttt{P1} and
\texttt{P2} by extending from the generated abstract React (EFSM state)
components and registering those to the session runtime component.

For the sake of code reuse, \cite{NoughtsAndCrosses}
uses \textit{higher-order components} (HOC) to build the correct state
implementations depending on which player the user chooses to be.
We leverage the \textit{Redux} \cite{Redux} state management library to keep
track of the game state, thus showing the flexibility of our library design in
being interoperable with other libraries and idiomatic JavaScript practices.
% REVIEW: tight-coupling?
Our approach encourages the separation of concerns between the
communication logic and program logic --- the generated session runtime
keeps track of the state of the EFSM to ensure protocol conformance by
construction, whilst \textit{Redux} solely manages our game state.