### 2020-03-09 PLACES_2020/related.tex

\section{Related Work}
The two main approaches for incorporating our MPST workflow into application
development are native language support for first-class linear channel
resources \cite{ATS} and code generation.
The latter closely relates to our proposal;
we highlight two areas of existing work under this approach that motivate our
design choice.

\paragraph{Endpoint API Generation}
Neykova and Yoshida targeted Python applications and the generation of runtime
monitors \cite{Python2017} to dynamically verify communication patterns.
Whilst the same approach could be applied to JavaScript, we can provide more
static guarantees with TypeScript's gradual typing system.
Scribble-Java \cite{Hybrid2016} proposed to encode the EFSM
states and transitions as classes and instance methods respectively, with
behavioural typing achieved statically by the type system and channel linearity
guarantees achieved dynamically since channels are exposed and
aliasing is not monitored.
% REVIEW: could the approach of the present paper by applied to Java as well? discuss?
Scribble-Java can generate callback-style APIs similar to the approach we
present, but this approach is arguably less idiomatic for Java developers.

\paragraph{Session Types in Web Development}
King et al.\ \cite{PureScript2019} targeted web development in PureScript using the
\textit{Concur UI} framework and proposed a type-level encoding of EFSMs as
multi-parameter type classes.
However, it presents a trade-off between achieving static linearity guarantees
from the type-level EFSM encoding under the expressive type system and
providing an intuitive development experience to developers, especially given
the prevalence of JavaScript and TypeScript applications in industry.
Fowler \cite{MVU2019} focused on applying binary session types in front-end web
development and presented approaches that tackle the challenge of guaranteeing
linearity in the event-driven environment, whereas our work is applicable to
multiparty scenarios.

Our work applies the aforementioned approaches in a \textit{multiparty} context
using industrial tools and practices to ultimately encourage MPST-safe web
application development workflows in industry.

%\paragraph{Typestate programming} \dots