### 2020-04-02 notes/notes.tex
Notes: init TeX

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{listings}

\title{Project Notes}
\author{Anson Miu}

\begin{document}

\nocite{*}

\maketitle

\section{Progress}

\paragraph{01/04/2020}
\begin{itemize}
\item 
Set up Node and TypeScript sandbox for testing.
The idea is to have a workflow where the
tool generates code for some protocol,
then copies into the sandbox and runs
the TypeScript compiler with the predefined
\texttt{tsconfig.json}.

\item
This guarantees that ``correct'' TypeScript is
generated with respect to the TypeScript 
language, not about correctness with respect
to session.
\end{itemize}

\paragraph{02/04/2020}
\begin{itemize}
\item
Set up React TypeScript sandbox for testing.
No need \texttt{create-react-app} at
the moment (which creates full application
boilerplate) as we are just trying to check
for TypeScript syntax.

\item
Piped the \texttt{tsfmt} into the code 
generation workflow so no need to deal with
spacing issues on Jinja template -- timesaver.

\item
The full code generation workflow ``works'' for 
both targets. The validity of the generated 
code works only for Node targets.

\item
Now focusing on formalising the React generated 
code -- taking more time than planned because
the TicTacToe example made more assumptions
about the protocol than I thought
\end{itemize}

\section{MPST Encoding}

\section{Code generation workflow}

\section{Testing}

\subsection{Unit tests}
TODO -- typical, use \texttt{unittest} module
to check for functionality of EFSM parsing
etc.

\subsection{Integration tests}
For a given protocol and specified role/target,
generate the TypeScript code, put into the
sandbox environment and run \texttt{npm test} 
(which is just a wrapper for 
\texttt{tsc --noEmit})
to expect success return code.

\section{To fix later (most important first)}

\subsection*{Receive transitions in React}
This is the current workflow -- need to verify whether the receive transition is safe?

\begin{enumerate}
\item
Assume you are on send state, user triggers
event to perform send.
\item
Runtime performs the send over the WebSocket
and renders the receive state \textbf{by
calling \texttt{setState}}. Calling setState 
schedules the state transition, just like
how a Promise resolve scheduels the call in
the high-priority callback queue.

\item When the receive state is rendered (as a 
result of \texttt{setState}), the lifecycle
method \texttt{componentDidMount} will call \texttt{register} to add the receive handlers

\item \textbf{At the same time}, message is sent from
other role to this user -- this triggers the 
onMessage event listener on the WebSocket. 
\textbf{There is no guarantee that \texttt{setState} 
is called before the message event listener.}

\end{enumerate}

If the onMessage event listener
is called before the setState,
then the implementation breaks because there is no
handler to use.

\subsubsection*{Possible solution}

Use the same approach as the backend.
\textit{Ignoring the exact TypeScript types for now,}
keep a message queue and a handler queue.

On mounting the receive state (which is the actual
transition), check whether there is a message in the 
queue.
\begin{itemize}
\item
If yes -- pop message and call handler based on label
(and cast payload accordingly).

\item
If no -- add handler to queue.
\end{itemize}

On message receive, check whether there are handlers
in the queue. 
\begin{itemize}
\item
If yes -- call the correct one based on
the label (and cast payload accordingly).

\item
If no -- add message to queue.
\end{itemize}

\begin{itemize}
\item
Do we need a queue? Is it possible for more than
one message to be queued up, since you advance to
a receive state and only expect to receive exactly
one message?

\item
Could use an optional-type variable
but need to consider the types...
\end{itemize}

\subsection*{Send transition in React}
Sends can only be event-driven --
currently not possible to do this:

\begin{enumerate}
\item Server sends browser 
\texttt{Offer(int)}
\item Browser immediately sends 
\texttt{Reject()} if payload is less than 0;
otherwise attach a \texttt{Counter(int)} 
send action to a user event.
\end{enumerate}

\section{Extensions}

\subsection*{Explicit connection actions}

See \cite{FASE2017} -- also refer to Battleships 
example from Jonathan.

\begin{lstlisting}
explicit global protocol Game(role Svr, role P1, role P2) {
  Play() connect P1 to Svr;
  Play() connect P2 to Svr;
  do OneGame(Svr, P1, P2);
  disconnect P1 and Svr;
  disconnect P2 and Svr;
 }

global protocol OneGame(role Svr, role P1, role P2)
  Pos(Point) from P1 to Svr;
  choice at Svr {
    Lose(Point) from Svr to P2;
    Win(Point) from Svr to P1;
  } or {
    Draw(Point) from Svr to P2;
    Draw(Point) from Svr to P1;
  } or {
    Update(Point) from Svr to P2;
    Update(Point) from Svr to P1;
    do OneGame(Svr, P2, P1);
  }
}

\end{lstlisting}

\bibliographystyle{acm}
\bibliography{notes}

\end{document}