### 2020-04-03 notes/notes.tex
Notes: update 03/04/2020

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{listings}

\title{Project Notes}
\author{Anson Miu}

\begin{document}

\maketitle

\section{Progress}

\paragraph{01/04/2020}
\begin{itemize}
\item 
Set up Node and TypeScript sandbox for testing.
The idea is to have a workflow where the
tool generates code for some protocol,
then copies into the sandbox and runs
the TypeScript compiler with the predefined
\texttt{tsconfig.json}.

\item
This guarantees that ``correct'' TypeScript is
generated with respect to the TypeScript 
language, not about correctness with respect
to session.
\end{itemize}

\paragraph{02/04/2020}
\begin{itemize}
\item
Set up React TypeScript sandbox for testing.
No need \texttt{create-react-app} at
the moment (which creates full application
boilerplate) as we are just trying to check
for TypeScript syntax.

\item
Piped the \texttt{tsfmt} into the code 
generation workflow so no need to deal with
spacing issues on Jinja template -- timesaver.

\item
The full code generation workflow ``works'' for 
both targets. The validity of the generated 
code works only for Node targets.

\item
Now focusing on formalising the React generated 
code -- taking more time than planned because
the TicTacToe example made more assumptions
about the protocol than I thought.
\end{itemize}

\paragraph{03/04/2020}
\begin{itemize}
\item
Build system testing pipeline -- register which
Scribble protocols to test, specify which role
(e.g. Bob) corresponds with which target (e.g. Node),
then the pipeline will perform the code generation,
check for success exit code, then invoke the
TypeScript compiler to check the TypeScript code
and also check for success exit code.
\end{itemize}

\section{MPST Encoding}

\subsection{Node}

TODO

\subsection{Browser}

TODO

\section{Code generation workflow}

\section{Testing}

\subsection{Unit tests}
TODO -- typical, use \texttt{unittest} module
to check for functionality of EFSM parsing
etc.

\subsection{Integration tests}
For a given protocol and specified role/target,
generate the TypeScript code, put into the
sandbox environment and run \texttt{npm test} 
(which is just a wrapper for 
\texttt{tsc --noEmit})
to expect success return code.

\texttt{npm test} is already set up -- just need to
set up the Python script to call for testing.

\section{To fix later (most important first)}

\subsection*{Receive transitions in React}
This is the current workflow -- need to verify whether the receive transition is safe?

\begin{enumerate}
\item
Assume you are on send state, user triggers
event to perform send.
\item
Runtime performs the send over the WebSocket
and renders the receive state \textbf{by
calling \texttt{setState}}. Calling setState 
schedules the state transition, just like
how a Promise resolve scheduels the call in
the high-priority callback queue.

\item When the receive state is rendered (as a 
result of \texttt{setState}), the lifecycle
method \texttt{componentDidMount} will call \texttt{register} to add the receive handlers

\item \textbf{At the same time}, message is sent from
other role to this user -- this triggers the 
onMessage event listener on the WebSocket. 
\textbf{There is no guarantee that \texttt{setState} 
is called before the message event listener.}

\end{enumerate}

If the onMessage event listener
is called before the setState,
then the implementation breaks because there is no
handler to use.

\subsubsection*{Possible solution}

Use the same approach as the backend.
\textit{Ignoring the exact TypeScript types for now,}
keep a message queue and a handler queue.

On mounting the receive state (which is the actual
transition), check whether there is a message in the 
queue.
If yes -- pop message and call handler based on label
(and cast payload accordingly).
If no -- add handler to queue.

On message receive, check whether there are handlers
in the queue. 
If yes -- call the correct one based on
the label (and cast payload accordingly).
If no -- add message to queue.

\begin{itemize}
\item
Do we need a queue? Is it possible for more than
one message to be queued up, since you advance to
a receive state and only expect to receive exactly
one message?

\item
Could use an optional-type variable
but need to consider the types...
\end{itemize}

\subsection*{Exception handling}

TODO read Exceptional GV paper \cite{ExceptionalGV}.

\subsection*{Send transition in React}
Sends can only be event-driven --
currently not possible to do this:

\begin{enumerate}
\item Server sends browser 
\texttt{Offer(int)}
\item Browser immediately sends 
\texttt{Reject()} if payload is less than 0;
otherwise attach a \texttt{Counter(int)} 
send action to a user event.
\end{enumerate}

\section{Extensions}

\subsection*{Explicit connection actions}

See \cite{FASE2017} -- also refer to Battleships 
example from Jonathan.

\begin{lstlisting}
explicit global protocol Game(role Svr, role P1, role P2) {
  Play() connect P1 to Svr;
  Play() connect P2 to Svr;
  do OneGame(Svr, P1, P2);
  disconnect P1 and Svr;
  disconnect P2 and Svr;
 }

global protocol OneGame(role Svr, role P1, role P2)
  Pos(Point) from P1 to Svr;
  choice at Svr {
    Lose(Point) from Svr to P2;
    Win(Point) from Svr to P1;
  } or {
    Draw(Point) from Svr to P2;
    Draw(Point) from Svr to P1;
  } or {
    Update(Point) from Svr to P2;
    Update(Point) from Svr to P1;
    do OneGame(Svr, P2, P1);
  }
}

\end{lstlisting}

\subsection*{Typestate in TypeScript}

Look into TyoeScript transfomers -- the idea is
to define typestate decorators, then augment the
TypeScript compiler to transform the code into some
intermediate layer to perform validation on
typestate: for example, consider the example of a 
stack that can either be full or empty.

\section{Report Structure}

\begin{enumerate}
\item 
Introduction -- as per PLACES (for now)

\item
Background -- process calculus, BST, MPST, code 
generation, TypeScript; think about where React goes?

\item
MPST encoding in TypeScript (section 2 in these notes).
Split into Node and browser targets, also talk about
the different ideas building up to this (pros and
cons of the experiments I did with Calculator in
2019).

Any of the new concepts (exception handling, etc.) will 
go in this section too.

TODO think about how to split into 
more logical sections.

\item
Code generation workflow (section 3 in these notes)
-- software architecture of
toolchain implemented in Python.

\end{enumerate}

\nocite{*}
\bibliographystyle{acm}
\bibliography{notes}

\end{document}