### 2020-05-25 report/sections/evalconcl/concl/contributions.tex
Report: detailed scaffolding

\section{Contributions}

\paragraph{Note: from PLACES}

We have presented an MPST-based framework for developing full-stack interactive
TypeScript applications with WebSocket communications.
The implementation conforms to a specified
protocol, statically providing linear channel usage guarantees and affine
channel usage guarantees for back-end and front-end targets respectively.