### 2020-05-25 report/sections/evalconcl/concl/future.tex

\section{Future Work}

\paragraph{Note: from PLACES}
Future work includes incorporating \textit{explicit connection actions}
introduced in \cite{ExplicitConnections} in our API generation to better model
real-world communication protocols that may feature in interactive web
applications.
% REVIEW: asynchronous
Server-side implementations may perform asynchronous operations on the
received messages, so supporting asynchronous values (such as JavaScript
\textit{Promises} \cite{promise}) in IO actions would be a welcome addition.
Whilst our approach supports multiparty sessions, the nature of
WebSockets require some server-based role in the communication protocol and
clients to interact via the server.
Extending support to WebRTC \cite{WebRTC} would
cater for peer-to-peer communication between browsers, which further opens up
possibilities for communication protocols supported by our approach.