### 2020-05-25 report/sections/server/codegen.tex

\chapter{\fancyname{SessionTS}: TypeScript API Generation}

\input{sections/server/codegen/workflow}

\input{sections/server/codegen/scribble}

\input{sections/server/codegen/efsm}

\input{sections/server/codegen/apigen}