### 2020-05-25 report/sections/server/node.tex

\chapter{\fancyname{NodeMPST}: Back-End Session Type Web Development}

\section{Motivation}

\section{Approach}

\section{State Encodings}

\subsection{Send}

\subsection{Receive}

\subsection{Terminal}

\section{EFSM Execution Runtime}

\section{Alternative Designs and Limitations}

\section{Summary}