### 2020-05-25 report/sections/general/implementation.tex
Report:  theory section scaffolding

\chapter{Implementation}

\section{Extending \fancyname{NodeMPST}}

\section{Extending \fancyname{ReactMPST}}

\section{Limitations}