### 2020-05-25 report/sections/general/theory.tex

\chapter{Theory: Routed Multiparty Session Types}

\section{Syntax}

\subsection{Global Types}

\subsection{Local Types}

\section{Labelled Transition System (LTS) Semantics}

\subsection{Trace Equivalence}

\subsection{Deadlock Freedom}

\section{Router-Parameterised Encoding}

\subsection{Preserving Well-formedness}

\subsection{Preserving Communication}