### 2020-05-28 report/sections/server/codegen.tex
Report: notes on part 1

\chapter{\fancyname{SessionTS}: Session Type API Generation for TypeScript}
\label{chap:codegen}

\input{sections/server/codegen/workflow}

\input{sections/server/codegen/scribble}

\input{sections/server/codegen/efsm}

\input{sections/server/codegen/apigen}

\input{sections/server/codegen/testing}