### 2020-05-28 report/sections/server/codegen/testing.tex

\section{Testing}
\begin{itemize}
\item unit tests for code generator executable
\item systems tests for type checking the generated output with the typescript compiler
\item sandbox environment set up on the docker container to set up different tsconfig.json files for both react/node targets
\end{itemize}