### 2020-05-28 report/sections/server/extensions/error.tex

\section{Error Handling}

\subsection{Motivation}
\begin{itemize}
\item existing work extends theory based on canonical session cancellations are from exceptions raised by logic
\item refer to 2FA example in Fowler paper on exceptional session types
\item need to also consider cancellations in a web-based scenario -- server dead, client closing their browser
\item need to also offer developer a seam for handling cancellations -- this needs to be well-parameterised so the developer knows how to handle the cancellation (for example, which role cancelled -- e.g. if it is a multi-way transaction, the server needs to know which role cancelled, so they can update the database accordingly).
\end{itemize}

\subsection{API Extension}
\begin{itemize}
\item developers just need to provide cancellation handlers along with the initial state
\item we standardise on parameterising it with role and reason: the reason is typed `any' to give developers the autonomy for implementation details
\end{itemize}

\subsection{Runtime Extension}
\begin{itemize}
\item build upon websocket RFC which specifies the `base' closing (or exit) code
\item we additionally define contracts for exit codes in our use case (e.g. role occupied, logical error), and make sure it does not override any websocket `traditions' as per the RFC
\item all based on well-defined event listeners on the websocket close and error events
\item must consider cancellation in two phases
\end{itemize}

\subsubsection{Cancellations during Session Joining}

\begin{itemize}
\item server -- emits role occupied
\item client -- closes browser whilst waiting for others
\end{itemize}

\subsubsection{Cancellations during EFSM Execution}

\begin{itemize}
\item server -- dies
\item server -- logical error
\item client -- dies
\item client -- logical error
\end{itemize}

\subsection{Limitations}
\begin{itemize}
\item Cancellation handlers are not typed -- namely, we use `any';
\item exception handler blocks in typescript are not typed
\item the `logical error' idea is also very vague as all code can be effectful in typescript (statically); cannot specifically say that the implementation is expected to `send payload of type int, but possibly throw error of type Error', which would otherwise be possible in other languages(?)
\end{itemize}