### 2020-05-28 report/sections/general/implementation.tex
Report: notes on part 2

\chapter{Implementation}

\begin{itemize}
\item key property from the theory is that the server's actions are effectively transparent, so we reflect that in the implementation
\end{itemize}

\section{Extending \fancyname{NodeMPST}}
\begin{itemize}
\item all messages have a recipient role
\item for messages not intended for server, simply route transparently
\end{itemize}

\section{Extending \fancyname{ReactMPST}}
\begin{itemize}
\item receive as before, but determine the sender from the payload since it could have originated from a non-server role
\item the double-queue system required for sending and receiving now need to be a map for React targets because the ordering is not preserved between roles (give example: A to B. C to B, server can route C's message before A's but B has to process A's first by protocol).
\end{itemize}