### 2020-05-28 report/sections/general/theory.tex

\chapter{Theory: Routed Session Types}

\section{Syntax}

\subsection{Local Types}

\begin{itemize}
\item routing construct
\item proxy send/receive
\end{itemize}

\subsection{Global Types}

\begin{itemize}
\item global type syntax
\item go over `routed communication'
\item participants
\item `acts as server role' (the diamond)
\item projection onto local types
\item well-formedness
\end{itemize}

\section{Labelled Transition System (LTS) Semantics}

\begin{itemize}
\item extend action syntax
\end{itemize}

\subsection{LTS over Global Types}

\subsection{LTS over Local Types}

\subsection{LTS over Collections of Local Types}

\subsection{Trace Equivalence}

\subsection{Deadlock Freedom}

\section{Router-Parameterised Encoding}

\begin{itemize}
\item define encoding on global types
\item define encoding on local types
\item define encoding on LTS actions
\item motivate with twobuyer example
\end{itemize}

\subsection{Correspondence between Encoding on Global and Local Types}
\begin{itemize}
\item prove proposition
\item state corollary for subsequent proof
\end{itemize}

\subsection{Preserving Well-formedness}

\subsection{Preserving Communication}