### 2020-05-29 report/sections/introduction/objectives.tex
Report: contributions

\section{Objectives}

The objective of this project is to \dots