### 2020-05-31 report/sections/server/codegen.tex
Report: start chapter 3

\chapter{\fancyname{SessionTS}: Session Type API Generation for TypeScript}
\label{chap:codegen}

In this chapter, we present
\fancyname{SessionTS}, an API generation toolchain for building
web applications in TypeScript that conform to a communication protocol 
based on multiparty session type theory.

\fancyname{SessionTS} supports communication protocols that define 
\emph{server-centric} topologies, meaning that: 
\textbf{(1)} there is exactly one participant executed on the 
Node.js runtime; 
\textbf{(2)} all other participants run on the web browser; and
\textbf{(3)} all non-server participants only communicate with the server.
We relax this assumption in \cref{part:general}.

The toolchain is publicly available at \cite{repo}.

\input{sections/server/codegen/workflow}

\input{sections/server/codegen/scribble}

\input{sections/server/codegen/efsm}

\input{sections/server/codegen/apigen}

\input{sections/server/codegen/testing}