### 2020-06-01 report/sections/general/p2p.tex
Report: start chapter 7

\chapter{Motivation: Supporting Peer-to-Peer Interactions}
\label{chap:p2p}

\section{\fancyname{Two Buyer} Protocol}

We present the \tprotocol{Two Buyer} protocol in \cref{lst:twobuyer}.

\begin{figure}[!ht]
\begin{lstlisting}[language=Scribble, tabsize=2]
module ClientClient;

global protocol TwoBuyer(role A, role B, role S) {
	title(string)      from A to S;
	quote(number)      from S to A;
	quote(number)      from S to B;
	quoteByTwo(number) from A to B;
	choice at B {
		ok()           from B to S;
		date(string)   from S to B;
	} or {
		quit()         from B to S;
	}
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Two Buyer} Protocol}
\label{lst:twobuyer}
\end{figure}

Assume \trole{S} is the server, and roles \trole{A}, \trole{B} run on the 
browser:

\begin{enumerate}
\item \trole{A} sends to \trole{S} the title of an item to sell;
\item \trole{S} sends a quote for said item to \trole{A} and \trole{B};
\item \trole{A} privately sends a separate quote to \trole{B};
\item \trole{B} makes a choice:
\begin{itemize}
\item if \trole{B} prefers the offer from \trole{S}, 
\trole{B} replies \tmsg{ok()} to \trole{S}, and \trole{S} informs
\trole{B} of the delivery date;

\item otherwise, \trole{B} takes the offer from \trole{A} and simply
responds \tmsg{quit()} to \trole{S}.
\end{itemize} 
\end{enumerate}

\begin{itemize}
\item TwoBuyer example -- highlight the sections that show the `arbitrary topology'
\item Not possible with WebSocket topology
\end{itemize}

\section{Server as a Router}
\begin{itemize}
\item How to make TwoBuyer example work just with WebSockets
\item Quickly highlight WebRtC as an option, but it has its own proprietary connection `protocol', and will still need to rely on a centralised server role to set up the peer to peer connection; including the setup in the code generation will introduce unnecessary complications in error handling which deviates from the focus of implementing typesafe web services
\item Essentially `decomposing' multiparty session into binary interactions with respect to the server
\end{itemize}

\section{Challenges}
\begin{itemize}
\item Routed communication is different from `normal' communication
- server cannot overserialise interactions
\item Existing work in decomposing multiparty sessions (see LinearDecompScala paper and the binary one it references) assume the `server' takes on the pure role as a router -- our work argues that the server can still participate in the multiparty session and concurrently route messages between clients in a way that does not overserialise the interactions and allow the routing to be transparent to other roles.
\item theory side -- need to prove that it is indeed possible for the server to carry out its own interactions and perform routing in a transparent way, as if the messages were directly peer to peer
\item implementation side -- changes need to reflect the theory and be as transparent to the developer as possible
\end{itemize}