### 2020-06-02 report/sections/general/p2p.tex
Report: progress on chapter 7

\chapter{Motivation: Supporting Peer-to-Peer Interactions}
\label{chap:p2p}

In this chapter,
we demonstrate how to support peer-to-peer interactions
between web browser endpoints over WebSocket transport,
thus relaxing the server-centric topology assumption from 
\cref{part:server}.

We motivate the problem using the 
\tprotocol{Two Buyer} protocol (\cref{section:twobuyer}),
outline our approach (\cref{section:routerapproach}),
and discuss the challenges we need to address 
(\cref{section:routerchallenges}) in order to validate the
correctness of our approach.

\section{\tprotocol{Two Buyer} Protocol}
\label{section:twobuyer}

We present the \tprotocol{Two Buyer} protocol in \cref{lst:twobuyer},
which describes a protocol between two buyers \trole{A}, \trole{B}
and a \trole{S}eller. We assume that the \trole{S}eller runs on
the server, and the buyers run on the browser.

\begin{figure}[!ht]
\begin{lstlisting}[language=scribble]
module ClientClient;

global protocol TwoBuyer(role A, role B, role S) {
	title(string)   from A to S;
	quote(number)   from S to A;
	quote(number)   from S to B;
	split(number)   from A to B; (*@\label{line:twobuyersplit}@*)
	choice at B {
		accept()    from B to A; (*@\label{line:twobuyerok}@*)
		buy()       from A to S;
	} or {
		reject()    from B to A; (*@\label{line:twobuyerno}@*)
		cancel()    from A to S;
	}
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Two Buyer} Protocol}
\label{lst:twobuyer}
\end{figure}

In the protocol, \trole{A} asks the \trole{S}eller for the quote
of a book \tmsg{title}. \trole{S}eller sends the \tmsg{quote}
to both buyers \trole{A} and \trole{B}. \trole{A} privately asks
proposes a \tmsg{split} with \trole{B}. Now, \trole{B} must make
a choice: 
either \trole{B} can \tmsg{accept} the split, where \trole{A} confirms
to \tmsg{buy} from the \trole{S}eller;
otherwise, \trole{B} will \tmsg{reject} the split, which
\tmsg{cancel}s the purchase.

\cref{line:twobuyersplit,line:twobuyerok,line:twobuyerno}
illustrate that the \tprotocol{Two Buyer} protocol 
does not implement a server-centric communication topology,
since these lines reference communication actions between
the two client browsers. 
As a result, \fancyname{SessionTS} cannot correctly generate
APIs that support the specified interactions 
for buyers \trole{A} and \trole{B}.

\section{Proposal: Server as a Router}
\label{section:routerapproach}

We acknowledge that peer-to-peer browser interactions
could be achieved via different transport abstractions, 
such as the WebRTC framework \cite{WebRTC}. 
However, WebRTC has its own proprietary ``protocol'' 
for establishing connections and exchanging network information,
which isn't formalised in session type theory. This setup phase
generally requires a \textit{signalling server} \cite{WebRTCSignaling},
which brings our problem full circle, as it can be implemented
over WebSocket transport. Extending API generation to also include
the boilerplate for initialising WebRTC connections could be an option,
but would introduce complications in the session runtime and
error handling (e.g. how to handle cancellations during the connection
phase?).
We believe that adopting WebRTC deviates from the focus of project in 
implementing type-safe webservices.

Hence, for the purpose of our work in this part,
we establish that WebSocket transport is an invariant, 
and proceed to build up our approach. 
We start by breaking down the concept of 
``peer-to-peer communication''.
Focusing on \cref{lst:twobuyer} \cref{line:twobuyersplit},
we can say that the interaction satisfies three properties:

\begin{enumerate}
\item 
As far as \trole{A} is concerned, \trole{A} is sending
a message that will be received by \trole{B}.

\item
As far as \trole{B} is concerned, \trole{B} is receiving
a message that was sent by \trole{A}.

\item
The \trole{S}eller is not involved in this interaction.
\end{enumerate}

We proceed to relax point 3: for the \trole{S}eller to be
not involved in the interaction, it suffices to guarantee that
the \trole{S}eller cannot intervene or hijack the communication.
This allows \trole{S}eller to ``oversee'' the communication.

On this basis, we motivate our approach to empower the 
server endpoint to act as a \textit{router}. 
That is to say, if the server endpoint receives a message
of which it isn't the intended recipient for, 
it routes the message to the intended recipient. 
Based on the WebSocket transport invariant,
all browser endpoints must join the session
through the server. This means that the server has information about
all browser endpoints, 
and is capable to perform the routing.

Returning to the \tprotocol{Two Buyer} example,
\trole{A} sends the \tmsg{split} message through its existing
WebSocket connection (with \trole{S}eller on the other side),
noting that \trole{B} is the intended recipient;
\trole{S}eller receives this through the WebSocket bound to
\trole{A}, and routes the message to \trole{B}, who receives it through
its existing connection with \trole{S}eller.
It is trivial that the three properties listed previously are still
satisfied: considering the relaxed version of point 3,
the semantics of routing a message corresponds to that of
overseeing the communication, in the sense that \trole{S}eller 
does not tamper with the payload or drop the message, so it is simply
overseeing the message being passed through.

\section{Challenges}
\label{section:routerchallenges}

% routing needs to be treated very carefully in the theory,
% we want to preserve communication, so the routed constructs must not % overserialise
% example: A->B: M1; S->B: M2.
% a valid trace is SB!M2, AB!M1;
% 
% if we treat routing messages with the same constructs,
% we get A->S: M1; S->B:M1; S->B:M2;
% then S
%}

\begin{itemize}
\item Routed communication is different from `normal' communication
- server cannot overserialise interactions
\item Existing work in decomposing multiparty sessions (see LinearDecompScala paper and the binary one it references) assume the `server' takes on the pure role as a router -- our work argues that the server can still participate in the multiparty session and concurrently route messages between clients in a way that does not overserialise the interactions and allow the routing to be transparent to other roles.
\item theory side -- need to prove that it is indeed possible for the server to carry out its own interactions and perform routing in a transparent way, as if the messages were directly peer to peer
\item implementation side -- changes need to reflect the theory and be as transparent to the developer as possible
\end{itemize}