### 2020-06-02 report/sections/general/p2p.tex
Report: restructure chapter 7

\chapter{Motivation: Supporting Peer-to-Peer Interactions}
\label{chap:p2p}

In this chapter,
we demonstrate how to support peer-to-peer interactions
between web browser endpoints over WebSocket transport,
thus relaxing the server-centric topology assumption from 
\cref{part:server}.

We motivate the problem using the 
\tprotocol{Two Buyer} protocol (\cref{section:twobuyer}),
outline our approach (\cref{section:routerapproach}),
and discuss the challenges we need to address 
(\cref{section:routerchallenges}) in order to validate the
correctness of our approach.

\input{sections/general/p2p/twobuyer}

\input{sections/general/p2p/approach}

\input{sections/general/p2p/challenges}