### 2020-06-03 report/sections/server/node.tex
Report: reorganise directory for chapter 4

\chapter{\fancyname{NodeMPST}: Back-End Session Type Web Development}
\label{chap:node}

In this chapter we present \fancyname{NodeMPST},
our session type API generation strategy for 
server-side endpoints implemented on the Node.js runtime \cite{node}.
We continue to use the running example of the \tprotocol{Adder}
protocol, and refer to the FSM of the \trole{Svr} endpoint
(\cref{fig:addersvrfsm}) throughout this chapter.

\begin{figure}[!b]
\centering
\includegraphics[width=0.5\textwidth]{AdderSvrFSM}
\captionof{figure}{\trole{Svr} Endpoint FSM
in \tprotocol{Adder} protocol}
\label{fig:addersvrfsm}
\end{figure}

\input{sections/server/node/challenges}
\input{sections/server/node/approach}
\input{sections/server/node/efsm}
\input{sections/server/node/runtime}
\input{sections/server/node/alt}
\input{sections/server/node/limitations}