### 2020-06-03 report/sections/server/node/alt.tex

\section{Alternative Designs}
\label{section:nodealt}

The main alternative EFSM encoding would be similar to those
presented in \cite{Hybrid2016}, which encodes
each EFSM state as a separate class and expose communication APIs
(e.g. send and/or receive) that respect the permitted transitions at
that state.

\begin{lstlisting}[language=scribble]
global protocol OneAdder(role Client, role Svr) {
	NUM1(number) from Client to Svr;
	NUM2(number) from Client to Svr;
	SUM(number)  from Svr to Client;
}
\end{lstlisting}

\begin{figure}[!h]
\centering
\begin{subfigure}{\textwidth}
\begin{lstlisting}[language=javascript, tabsize=2]
const logic = async (init) => {
	const [x, num2] = await init.receive();
 	const [y, sum] = await num2.receive();
 	// init.receive();	// this would type-check but violate linearity
	return sum.send(x + y);
}
\end{lstlisting}
\caption{\dots}
%\label{...}
\end{subfigure}
\hfill
\begin{subfigure}{\textwidth}
\begin{lstlisting}[language=javascript]
const logic = new S4({
	NUM1: (x) => new S6({
		NUM2: (y) => ([Labels.S7.SUM, [x + y], new S5()]),
	}),
});
\end{lstlisting}
\caption{\dots}
%\label{...}
\end{subfigure}
\captionof{figure}{\dots}
%\label{...}
\end{figure}

\begin{itemize}
\item exposing channel resources like the java example -- less idiomatic (not-event-driven), need runtime linearity checks like the java example
\item `hardcoded' switch case runtime like the TSOP papers -- doesn't best leverage the type system
\end{itemize}