### 2020-06-03 report/sections/server/node/limitations.tex

\section{Limitations}
\label{section:nodelimitations}

\begin{itemize}
\item verbose
\item state identifiers are in the API
\item scoping achieved through nested objects
\item performance aspect with class-based handlers
\end{itemize}

\begin{lstlisting}[language=javascript]
const first = new S4({ NUM1: (x) => secondOperand, });
const second = new S6({ NUM2: (y) => sum, });
const sum = [Labels.S7.SUM, ???, new S5()];
\end{lstlisting}