### 2020-06-03 report/sections/general/p2p/twobuyer.tex
Report: minor change to TwoBuyer

\section{\tprotocol{Two Buyer} Protocol}
\label{section:twobuyer}

We present the \tprotocol{Two Buyer} protocol in \cref{lst:twobuyer},
which describes a protocol between two buyers \trole{A}, \trole{B}
and a \trole{S}eller. We assume that the \trole{S}eller runs on
the server, and the buyers run on the browser.

\begin{figure}[!ht]
\begin{lstlisting}[language=scribble]
global protocol TwoBuyer(role A, role B, role S) {
	title(string)   from A to S;
	quote(number)   from S to A;
	quote(number)   from S to B;
	split(number)   from A to B; (*@\label{line:twobuyersplit}@*)
	choice at B {
		accept()    from B to A; (*@\label{line:twobuyerok}@*)
		buy()       from A to S;
	} or {
		reject()    from B to A; (*@\label{line:twobuyerno}@*)
		cancel()    from A to S;
	}
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Two Buyer} Protocol}
\label{lst:twobuyer}
\end{figure}

In the protocol, \trole{A} asks the \trole{S}eller for the quote
of a book \tmsg{title}. \trole{S}eller sends the \tmsg{quote}
to both buyers \trole{A} and \trole{B}. \trole{A} privately asks
proposes a \tmsg{split} with \trole{B}. Now, \trole{B} must make
a choice: 
either \trole{B} can \tmsg{accept} the split, where \trole{A} confirms
to \tmsg{buy} from the \trole{S}eller;
otherwise, \trole{B} will \tmsg{reject} the split, which
\tmsg{cancel}s the purchase.

\cref{line:twobuyersplit,line:twobuyerok,line:twobuyerno}
illustrate that the \tprotocol{Two Buyer} protocol 
does not implement a server-centric communication topology,
since these lines reference communication actions between
the two client browsers. 
As a result, \fancyname{SessionTS} cannot correctly generate
APIs that support the specified interactions 
for buyers \trole{A} and \trole{B}.