### 2020-06-04 report/sections/server/node/runtime.tex
Report: progress on chapter 4 runtime

\section{Runtime}
\label{section:noderuntime}

We define the session runtime for the \trole{Svr} endpoint
of the \tprotocol{Adder} protocol in Svr.ts,
named after the endpoint.
It exposes a \textbf{public API}
(\cref{lst:noderuntimepublicapi})
with seams for the developer to pass in the WebSocket server 
and application logic
(i.e. their handler implementations).
It is the responsibility of the developer to construct the
WebSocket server and set it up to listen for incoming connections.
Internally, it keeps a \textit{private API}
for executing the EFSM, when all participants have 
joined the session.

The role of the public API is to manage incoming connections and
wait for all participants to join the session 
(\cref{subsection:noderuntimepublic}), before
handing off to the private API to execute the EFSM
(\cref{subsection:noderuntimeprivate}).

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,title=Svr.ts]
export class Svr {
	constructor(wss: WebSocket.Server,
				initialState: Implementation.S51) { ... }
	...
}

// Not exported to developer
class Session {
	private wss: WebSocket.Server;
	private initialState: Implementation.S51;
	private roleToSocket: RoleToSocket;
	...
}
\end{lstlisting}
\captionof{lstlisting}{Public and Private API of Session Runtime for Server Endpoint}
\label{lst:noderuntimepublicapi}
\end{figure}

\subsection{Managing Connections}
\label{subsection:noderuntimepublic}

\begin{itemize}
\item use set and partial object to keep track of pending connections
\item ws event listeners
\item forward reference that we need to manage cancellations here too
\end{itemize}

\subsection{Executing the EFSM}
\label{subsection:noderuntimeprivate}

The \texttt{Session} class executes the EFSM.
We define a transition function, \texttt{next()}, 
parameterised by the current state,
which invokes the handler defined by the develoepr
and performs the required channel actions for non-terminal states:

\begin{itemize}
\item 
For \textbf{send} states, the handler will
return the label and payload to be sent, along with the
successor state implementation. The transition function
should construct and send the message, 
and transition to the successor state.

\item
For \textbf{receive} states, we change the
message event listener on the WebSocket to
pass the incoming message to the handler.
The handler will return the successor state,
which the runtime can transition to.

\end{itemize}

We conceptualise this in \cref{lst:noderuntimesimple}.
The discriminated union lets the runtime figure out
the type of the current state.

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
next(state: Implementation.Type) {
	switch (state.type) {
		case 'Send': {
			const [label, payload, succ]: (*@\hl{???}@*) = state.handler;	 (*@\label{line:noderuntimebadtype1}@*)
			this.send((*@\hl{???}@*), label, payload); (*@\label{line:noderuntimeunknownrole}@*)		
			return this.next(succ);
		}
		case 'Receive': {
			this.wss.onmessage = ({ data }) => {
				const { label, payload } = JSON.parse(data) as (*@\hl{???}@*); (*@\label{line:noderuntimebadtype2}@*)
				const succ: (*@\hl{???}@*) = state.handler[label](...payload); (*@\label{line:noderuntimebadtype3}@*)
				return this.next(succ);
			}
		}
		case 'Terminate': { return; }
	}
}
\end{lstlisting}
\captionof{lstlisting}{Conceptual EFSM Transition Function 
for Server-Side Endpoint}
\label{lst:noderuntimesimple}
\end{figure}

However, we still face problems with resolving types, as highlighted.
Just because we know that the current state is a send state,
we do not know \textit{which} particular state it is,
so we cannot accurately type the handler (\cref{line:noderuntimebadtype1}).
The same problem is amplified for the receive state:
we need to know the specific receive state in order
to correctly serialise the message (\cref{line:noderuntimebadtype2})
and interpret the successor state (\cref{line:noderuntimebadtype3}).
We see another problem with handling send states:
because we do not know the specific send state, we do not know which
participant to send the message to (\cref{line:noderuntimeunknownrole}).

These all reduce to the same core problem: the runtime
needs to know the specific state at compile-time\footnote{
Whether TypeScript ``compiles'' or ``transpiles''
(or even \textit{``transcompiles''} \cite{transcompiles}) to JavaScript
is not relevant to our work; we stick with compilation
and keep our terminology consistent.}.
We solve this through \textit{runtime polymorphism} instead, since
the specific type of \texttt{state} is known at runtime.
For each type of state, we define a common API that can be invoked
by the EFSM transition function. To achieve runtime polymorphism,
each concrete state must provide a specific implementation: 
this motivates our design for defining the discriminated union
using abstract classes with abstract methods 
(\cref{lst:nodeabstractclass}).

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
abstract class ISend {
	type: 'Send' = 'Send';
	abstract performSend(
		next: EfsmTransitionHandler,
		sendMessage: (role: Roles.Peers, label: string, payload: any[])
			=> void,
	): void;
};

abstract class IReceive {
	type: 'Receive' = 'Receive';
	abstract prepareReceive(
		next: EfsmTransitionHandler,
		register: (from: Roles.Peers, messageHandler: MessageHandler)
			=> void,
	): void;
};
\end{lstlisting}
\captionof{lstlisting}{\dots}
\label{lst:nodeabstractclass}
\end{figure}

Using this approach,
we can pass the transition function and channel actions 
from the \texttt{Session} runtime class
to the individual state \texttt{Implementation} classes:
these are also generated by \fancyname{NodeTS}, so we
guarantee linear usage of channel resources by construction as well.
This \textit{significantly} simplifies the design of the runtime
(\cref{lst:noderuntime}),
because the transition function no longer needs to know the type 
of the specific state at compile-time.

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
next(state: Implementation.Type) {
	switch (state.type) {
		case 'Send':
			return state.performSend(this.next, this.send);
		case 'Receive':
			return state.prepareReceive(this.next, this.register);
		case 'Terminal':
			return;
	}
}
\end{lstlisting}
\captionof{lstlisting}{Final EFSM Transition Function for Server-Side
Endpoint}
\label{lst:noderuntime}
\end{figure}

By passing the specific functions defined in the runtime
to the individual EFSM states, we can visualise our runtime
as a \textit{message passing} abstraction 
(\cref{fig:noderuntimeefsm}): the runtime uses the common
\texttt{performSend()} or \texttt{prepareReceive()} API
to delegate to the specialised implementation, which will in turn
ask the runtime to perform specialised channel actions using
the parameterised methods, and finally delegate back to
the runtime to transition to the specific successor state.

We illustrate

\begin{itemize}
\item implementation details delegated to the actual implementation states
\item all non-terminal states will `return' the successor implementation -- very difficult to resolve in the type-checker in the runtime (give example of what it may look like) -- solve by giving each state implementation the `advance' function to respect the event-driven nature of everything, so they can call it on completion
\item visualise the interaction between runtime and states as a sequence diagram for message passing
\item constructor binds all methods (that will be passed to other components0 to `this' because of how javascript works with respect to the scoping of `this' (give short example -- or ignore, if this is in the background for typescript)
\item advance -- use discriminated union to figure out what to pass to the state (send -- sendMessage, receive -- register)
\end{itemize}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.5\textwidth]{NodeRuntimeEFSM}
\captionof{figure}{``Message Passing'' Abstraction of EFSM Execution for
Server Endpoints}
\label{fig:noderuntimeefsm}
\end{figure}

\subsection{Sending Messages}
Message Types are defined as interfaces, which
are represented by objects. By convention,
we serialise messages into \textit{JavaScript Object Notation}
(or JSON) \cite{json} using the built-in \texttt{JSON.stringify()}
method.

They are decoded using the same interface schema on the receiving end
using \texttt{JSON.parse()}: whilst the method return type is the
dynamic \lstonelinejs{any} type, 
we guarantee type safety by construction
as we performed the serialisation in the first place, so
we can safely interpret the deserialised content 
using a concrete type.

\subsection{Handling Message Receives}
\begin{itemize}
\item motivate the edge case that message can arrive in the websocket (in succession) before the handler is registered
\item motivate the edge case that message can arrive `out of protocol order' 
\item explain the double queue system and how that provides consistency -- emphasising that this works because of the single-threaded typescript runtime
\end{itemize}

\begin{figure}[!ht]
\centering
\begin{subfigure}[b]{0.8\textwidth}
\centering
\includegraphics[width=\textwidth]{NodeRuntimeReceive2}
\caption{Message processed before transitioning to receive state}
\label{subfig:nodereceivemsgfirst}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.8\textwidth}
\centering
\includegraphics[width=\textwidth]{NodeRuntimeReceive1}
\caption{Message processed after transitioning to receive state}
\label{subfig:nodereceivehandlefirst}
\end{subfigure}
\captionof{figure}{Possible Orderings for Handling Message Event and Preparing
Receive State}
\label{fig:nodereceivecompare}
\end{figure}

\subsection{Handling Termination}
WebSocket connections should be closed when the session terminates,
and both the browser endpoint and the server endpoint are capable
of closing connection. As we generate code for both, we define
a convention that the browser endpoint will initiate the WebSocket
close event.