### 2020-06-05 report/sections/server/react.tex
Report: restructure directory for React section

\chapter{\fancyname{ReactMPST}: Front-End Session Type Web Development}
\label{chap:react}

In this chapter, we
\dots

\input{sections/server/react/challenges}
\input{sections/server/react/approach}
\input{sections/server/react/efsm}
\input{sections/server/react/runtime}
\input{sections/server/react/alt}
\input{sections/server/react/limitations}

\section{Summary}
\begin{itemize}
\item static guarantees on avoiding communication mismatch
\item static guarantees on affine channel usage
\item leverage the react framework to express the model types approach from Fowler 2020 appropriately, to provide static guarantees, by construction, that the channel actions made available at each state respect the protocol, and by definition, multiparty session type safety
\end{itemize}