### 2020-06-05 report/sections/server/react/challenges.tex

\section{Challenges}
\begin{itemize}
\item challenges of session typing user interface logic -- how does one formalise channel linearity? (give example of how complicated this actually is -- e.g. how to guarantee that, if a button is triggered for sending, that the action will not be triggered at other states?)
\item existing works use non-mainstream language -- motivate Fowler LINKS, PureScript ConcurUI, identify that it needs some form of linear typing mechanism or `sequential building' UI to provide the safety properties
\item challenge -- provide the same static communication safety guarantees for front-end logic written in typescript, in a way that is usable and idiomatic for the average typescript developer?
\end{itemize}