### 2020-06-05 report/sections/server/react/limitations.tex

\section{Limitations}
\begin{itemize}
\item affine channel usage guarantees -- cannot statically guarantee that the factory is attached to an event visible on the DOM
\item decoupling communication state management from business logic state management -- runtime design does not permit passing non-communication-related data as props; need to use other methods for propagating applicaton data (such as React Context API or Redux)
\item the point above might not be a `limitation', but more of an encouragement to change practices because it makes sense, architecturally, to separate concerns
\end{itemize}