### 2020-06-05 report/sections/server/react.tex
Report: progress on chapter 5 challenges

\chapter{\fancyname{ReactMPST}: Front-End Session Type Web Development}
\label{chap:react}

In this chapter, we present \fancyname{ReactMPST},
our session type API generation strategy for browser-side endpoints
implemented using React \cite{React}.
We will refer to the FSM of the \trole{Client}
endpoint (\cref{fig:adderclientfsm})
from the \tprotocol{Adder} protocol 
throughout this chapter.

\begin{figure}[!b]
\centering
\includegraphics[width=0.55\textwidth]{AdderClientFSM}
\captionof{figure}{\trole{Client} Endpoint FSM
in \tprotocol{Adder} protocol}
\label{fig:adderclientfsm}
\end{figure}

\input{sections/server/react/challenges}
\input{sections/server/react/approach}
\input{sections/server/react/efsm}
\input{sections/server/react/runtime}
\input{sections/server/react/alt}
\input{sections/server/react/limitations}

\section{Summary}
\begin{itemize}
\item static guarantees on avoiding communication mismatch
\item static guarantees on affine channel usage
\item leverage the react framework to express the model types approach from Fowler 2020 appropriately, to provide static guarantees, by construction, that the channel actions made available at each state respect the protocol, and by definition, multiparty session type safety
\end{itemize}