### 2020-06-05 report/sections/server/react/alt.tex
Report: progress on chapter 5 alternative designs

\section{Alternative Designs}

For send states, a simpler approach would be
to provide the developer with a \texttt{send()} function 
for each permitted selection.
The factory approach would still apply -- the runtime could pass something
of the form

\begin{lstlisting}[language=javascript, numbers=none]
declare function buildSend<T>(label: string): (payload: T) => void
\end{lstlisting}

as a prop to the send state component.
By passing the payload type and label, the send state component
can build a handler (bound to the payload type) that performs a
send over the WebSocket when called.
This provides more flexibility for the developer's implementation,
but this clearly exposes channel resources and 
comes at the cost of not being able to provide guarantees
on affine channel usage.

For receive states, \dots

%
%\begin{itemize}
%\item not sure whether to put this in the subsection above so it flows better?
%\item naive implementation is to pass the send() function as a prop for the developer to incorporate in their event listener -- this is most flexible, but we cannot provide static guarantees
%\item factory approach -- the runtime could pass a factory function which, takes an UI element and performs the binding; this is better, as channel resources are not exposed, but the adversarial-minded developer can later override the event listener in the DOM etc
%\end{itemize}