### 2020-06-07 report/sections/abstract.tex
Report: progress on abstract

\begin{abstract}

We present an end-to-end framework
to write interactive web applications,
communicating over WebSocket transport,
that statically conform to prescribed communication protocols.

We implement a code generation toolchain to
provide developers with TypeScript APIs generated
from a communication protocol specification 
based on multiparty session type theory.

Our toolchain generates APIs that are compatible with
the Node.js runtime and the React.js framework
for server-side and browser-side endpoints respectively.

We evaluate the expressiveness of our API generation solution
using three case studies of protocols found in web-services,
and analyse the overhead of our implementation through 
running micro-benchmarks on a simple binary session protocol.

\end{abstract}