### 2020-06-07 report/sections/abstract.tex
Report: complete abstract

\begin{abstract}

\noindent
We present an end-to-end framework
to write interactive web applications
communicating over WebSocket transport
that conform to prescribed communication protocols.
We implement a code generation toolchain to
provide developers with TypeScript APIs generated
from a communication protocol specification 
based on multiparty session type theory.

Our work is compatible with modern web programming
industrial practices. The APIs
generated by the toolchain 
build upon TypeScript concurrency primitives,
complement the event-driven style of programming
in full-stack web development,
and are compatible
with the Node.js runtime and the React.js framework
for server-side and browser-side endpoints respectively.

We propose an approach for supporting peer-to-peer communication
between browser-side endpoints through routing communication via the server
in a way that avoids excessive serialisation
and preserves communication safety. 
We formalise this approach by extending multiparty session type theory
and proving that properties guaranteeing communication safety
are preserved.

We evaluate the expressiveness of our API generation solution
using three case studies of protocols found in web-services,
and analyse the overhead of our implementation through 
running micro-benchmarks on a simple binary session protocol.

\end{abstract}