### 2020-06-10 report/sections/evalconcl/concl/contributions.tex
Report: complete conclusion

\section{Contributions}

In this project, we develop and present an
MPST-based framework for developing full-stack
interactive TypeScript applications over WebSocket transport.

We motivate our API generation approach from 
\cite{Hybrid2016,PureScript2019} to generate TypeScript APIs
from a Scribble protocol specification.
Developers write their full-stack applications
by implementing the generated APIs, and can enjoy 
communication safety guarantees by construction.
We explain how our session type encodings target the Node.js
server-side runtime and React.js framework in
\cref{chap:node,chap:react} for server-side
and browser-side endpoints respectively.

In particular, we present a novel approach for integrating
session types into web-based GUI programming based on
translating the theory on model types \cite{MVU2020}
to idiomatic practices on the React.js framework.
With respect to session type theory, implementations using
the generated APIs statically enjoy linear channel usage
guarantees and affine channel usage guarantees for back-end
and front-end targets respectively.

Compared to previous work, we are not only able to
\textbf{statically} provide the same level of communication safety
guarantees from multiparty session type theory, but we do so
using modern web programming practices to increase
the relevance and usability of our work in industry:
our work targets TypeScript, Node.js and React.js,
and we explain in \cref{chap:ext} how we support advanced
web development idioms such as asynchronous implementations.

In addition, existing proposals on API generation for web development
over WebSockets \cite{PureScript2019} only support protocols
implementing server-centric topologies, in order to be 
compatible with WebSocket transport.
We formalise a theory of routed multiparty session types in 
\cref{chap:theory} to prove that it is possible to relax the
server-centric topology assumption over WebSocket transport in a way 
that preserves communication along with the communication safety 
properties inherited from canonical session type theory.

% PLACES
Our work has received positive feedback from academia.
The project has been accepted and published in the
Proceedings of the 12th International Workshop on
Programming Language Approaches to Concurrency- \& 
Communication-cEntric Software (PLACES 2020) \cite{PLACES2020}.
We were also invited to present this work as 
\emph{Generating Interactive WebSocket Applications in TypeScript}
at PLACES 2020.

% OVERALL
Overall, we offer an end-to-end solution for integrating
multiparty session type theory into modern full-stack web programming
practices, in a way that actively encourages developers to
design their application with communication safety in mind.
Our formalism of routed multiparty session types enables our
API generation tool to support a wider range of protocols, and
further reveals the potential
for implementing dynamic communication structures over centralised
network topologies.