### 2020-06-10 report/sections/evalconcl/eval.tex

\chapter{Evaluation}
\label{chap:eval}
%benchmark -- compile time / benchmarking

\section{Multiparty Sessions: \fancyname{Noughts and Crosses}}
\label{section:evalgame}
\begin{itemize}
\item sample client implementation -- show the flexibility of factory API for sending to be used in the UI part (since move selection can be be made )
\item show that not all the UI is being controlled by the runtime (i.e. calculator app has a header/fake navbar)
\item show the calculator logic achievable using React Context
\item sample server implementation + snippet of generated code (full in appendix)
\end{itemize}

\begin{figure}[!ht]
\begin{lstlisting}[language=Scribble, tabsize=2]
module NoughtsAndCrosses;
type <typescript> "Coordinate" from "./Types" as Point;

global protocol Game(role Svr, role P1, role P2) {
	Pos(Point) from P1 to Svr;
	choice at Svr {
		Lose(Point)   from Svr to P2;
		Win(Point)    from Svr to P1;
	} or {
		Draw(Point)   from Svr to P2;
		Draw(Point)   from Svr to P1;
	} or {
		Update(Point) from Svr to P2;
		Update(Point) from Svr to P1;
		do Game(Svr, P2, P1);
	}
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Noughts and Crosses} Protocol}
\label{lst:game}
\end{figure}

\section{Routed Multiparty Sessions: \fancyname{Two Buyers}}
\begin{itemize}
\item show that the routing aspect is transparent to both sides
\end{itemize}

\section{Delegated Sessions: \fancyname{ATM}}
\label{section:evalatm}

\begin{itemize}
\item Show that delegation is not supported
even if the protocol ``allows''
\end{itemize}

\begin{figure}[!ht]
\begin{lstlisting}[language=Scribble, tabsize=2]
module Delegation;

global protocol ATM(role Svr, role Client) {
	choice at Client {
		WITHDRAW(string, number) from Client to Svr;
		choice at Svr {
			BALANCE(number) from Svr to Client;		
		} or {
			FAIL(string) from Svr to Client;		
		}
		do ATM(Svr, Client);
	} or {
		DEPOSIT(string, number) from Client to Svr;
		BALANCE(number) from Svr to Client;
		do ATM(Svr, Client);
	} or {
		BYE(string) from Client to Svr;
		BYE() from Svr to Client;	
	}
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{ATM} Protocol}
\label{lst:atm}
\end{figure}

\input{sections/evalconcl/eval/benchmarks}