### 2020-06-10 report/sections/introduction/objectives.tex
Report: progress on introduction

\section{Objectives}

The objective of this project is to provide an end-to-end
framework for writing full-stack web applications that
\textbf{statically} conform to a communication protocol.
Developers specify their communication protocol using
the Scribble protocol description language \cite{Scribble}
which is based on multiparty session type (MPST) theory \cite{MPST}.
The Scribble protocol is used as a basis for generating
TypeScript APIs targeted for the Node.js server-side runtime
and React.js browser-side framework. 
By implementing the generated APIs in their full-stack applications,
developers can enjoy \textbf{static} guarantees with respect 
to communication safety: there will not be any communication mismatches,
and channel linearity will be respected.
Finally, we extend our framework to support communication
protocols that cannot be natively implemented over
a server-centric network topology using WebSockets.
This involves defining a theory of routed multiparty session types
to formalise the preservation of communication safety properties.