### 2020-06-10 report/sections/server/codegen.tex
Report: progress on chapter 6.1

\chapter{\fancyname{SessionTS}: Session Type API Generation for TypeScript}
\label{chap:codegen}

In this chapter, we present
\fancyname{SessionTS},
a toolchain for generating TypeScript APIs that developers can
use to write web applications that conform to their specified
communication protocol.
The toolchain is publicly available at \cite{repo}.

\fancyname{SessionTS} supports communication protocols that define 
\emph{server-centric} topologies, meaning that: 
\textbf{(1)} there is exactly one participant executed on the 
Node.js runtime; 
\textbf{(2)} all other participants run on the web browser; and
\textbf{(3)} all non-server participants only communicate with the server.
Examples of such protocols include the \tprotocol{Noughts and Crosses}
multiplayer game (\cref{section:evalgame}) and the \tprotocol{ATM} protocol.
We relax this assumption in \cref{part:general}.

\input{sections/server/codegen/workflow}

\input{sections/server/codegen/implementation}

\input{sections/server/codegen/testing}