### 2020-06-10 report/sections/server/extensions/async.tex

\section{Supporting Asynchronous Implementations}
\label{section:async}

\subsection{Motivation}

Consider a basic example of two-factor authentication 
(\tprotocol{2FA}) in \cref{lst:2FA} motivated from \cite{Exceptional}:
\trole{Client} tries to log in to \trole{Svr}, which either
authorises or denies the login attempt. If \trole{Client} is logging 
in from a new device, the \trole{Svr} sends a challenge key
and waits for a response code before proceeding.

\begin{figure}[!h]
\begin{lstlisting}[language=scribble]
global protocol 2FA(role Svr, role Client) {
	Login(string, string) from Client to Svr;
	choice at Svr { (*@\label{line:2fachoice1}@*)
		Authorised() from Svr to Client;
		do Main(Svr, Client);		// Details are not important
	} or {
		Challenge(string)	from Svr to Client;
		Response(number) 	from Client to Svr;
		choice at Svr { (*@\label{line:2fachoice2}@*)
			Authorised() from Svr to Client;
			do Main(Svr, Client);
		} or {
			AccessDenied() from Svr to Client;
		}
	} or {
		AccessDenied() from Svr to Client;	
	}
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Two Factor Authentication} Protocol}
\label{lst:2FA}
\end{figure}

The developer's implementation on 
\cref{line:2fachoice1,line:2fachoice2} is likely to query
a database for credential verification.
Database operations are implemented using \textit{asynchronous},
\textit{non-blocking} APIs; it is
recommended that file operations do not block the main execution,
so a JavaScript database API may require a \textit{callback}
to propagate the return value when it is available.
A server endpoint implementation of \cref{lst:2FA}
may resemble the following:

\begin{lstlisting}[language=javascript]
new Implementation.S40({
	Login: (username: string, password: string) => {
		db.lookup(username, password, (result) => {
			// returns Implementation.S42 here
		});
	}
});
\end{lstlisting}

Callbacks are commonplace in dealing with concurrency in JavaScript
\cite{CallbackHell}; we will go over other widely-used
concurrency primitives shortly.
Whatever the case may be, asynchronous developer implementations
using our generated APIs will not type-check as the function
returns immediately (hence, has return type \lstonelinejs{void}),
even though the callback function has the expected return type.
The developer will encounter the following:

\begin{lstlisting}[tabsize=2,numbers=none]
Type '(username: string, password: string) => void' is not 
assignable to type '(payload_0: string, payload_1: string) => S42'.
\end{lstlisting}

This is also possible for browser-side implementations:
when the user invokes a UI event, the browser-side logic
may first consult a third-party API for additional information
(which is not part of the communication protocol)
before performing a send action; the \textit{Fetch} API \cite{Fetch}
for making API calls is asynchronous.

In order to make \fancyname{SessionTS} more relevant and compatible
with modern web programming practices, it is important to
support asynchronous developer implementations as part of our
generated APIs and runtime.

This means we should support
the concurrency primitives built into TypeScript:

\subparagraph{Callbacks}
Callback-based APIs are higher-order functions -- they take functions
as parameters and invoke them when the asynchronous operation is complete.
The \texttt{setTimeout(fn, delay)} function available on the browser
waits for \texttt{delay} milliseconds before invoking \texttt{fn()} --
or more precisely, before queuing the execution of \texttt{fn()}.
As a result, the code listing below prints \lstonelinejs{'Second!'}
before \lstonelinejs{'First!'}, even though \texttt{setTimeout}
was called with 0ms delay.

\begin{lstlisting}[language=javascript,numbers=none]
setTimeout(() => console.log('First!', 0);
console.log('Second!');
\end{lstlisting}

\subparagraph{Promises}
A Promise formalises the ``completion'' callback
and its error handling construct. The developer passes a ``success''
handler and ``error'' handler, conventionally named
``then'' and ``catch'' respectively, to construct a Promise.
This is analogue to the \texttt{Future} interface in Java \cite{JavaFuture}.
The Promise will invoke the correct handler on completion.
The benefits of Promises in our context is that TypeScript allows
us to specify the type of payload that the Promise resolves to.

\begin{lstlisting}[language=javascript,numbers=none]
declare function promise(n: number): Promise<number>
const promise = (n: number) => 
	new Promise((resolve, reject) => {
		if (n < 10)  resolve(n));
		else reject('Number too big!');
	});

declare function handleNumber(n: number) { ... } 
declare function handleError(err: string) { ... }

promise(10).then(handleNumber).catch(handleError)
\end{lstlisting}

\subparagraph{Async/Await}
The \lstonelinejs{async}/\lstonelinejs{await} construct
is syntactic sugar for Promises. An \lstonelinejs{async}
function always returns a Promise, and marking \lstonelinejs{await}
for the identifier of the return value for a function returning a Promise
``unwraps'' the resolved payload. A downside is that error handling
reverts to the \lstonelinejs{try ... catch} construct.

\begin{lstlisting}[language=javascript,numbers=none]
const asyncSum = async (x: number, y: number) => {
	return (await promise(x)) + (await promise(y));
}

asyncSum(1, 2);  // returns Promise<number> that resolves to 3
\end{lstlisting}

\subsection{API Extension}
\begin{itemize}
\item should be very short section -- use union type MaybePromise so developer can provide either
\end{itemize}

\begin{lstlisting}[language=javascript,numbers=none]
type MaybePromise<T> = T | Promise<T>;
type FromPromise<T> = T extends MaybePromise<infer R> ? R : never;
\end{lstlisting}

\begin{lstlisting}[language=javascript,tabsize=2]
new Implementation.S40({
	Login: (*@\hl{async}@*) (username: string, password: string) => {
		const result = (*@\hl{await}@*) db.lookup(username, password);
		if (result.valid)
			return new Implementation.S42('Authorised', ...);
		else if (result.unauthorised)
			return new Implementation.S42('AccessDenied', ...);
		else
			return new Implementation.S42('Challenge', ..., challenger);
	}
});

const challenger = (key: number) => {
	return new (*@\hl{Promise}@*)((resolve, reject) => {
		keychain.check(key, ok => {
			if (ok)
				resolve(new Implementation.S44('Authorised', ...);
			else 
				resolve(new Implementation.S44('AccessDenied', ...);	
		});
	});
}
\end{lstlisting}

\subsection{Runtime Extension}

Generally, minimal changes are required for the runtime.
The key observation is that the concrete types for the any
handlers are the same -- they just might be wrapped in a Promise
construct.
To handle potential Promises, we wrap existing logic
as a delayed computation, colloquially known as a \textit{thunk}
\cite{Thunk}.

\begin{itemize}
\item server send actions -- use the Thunk pattern to wrap the existing logic, the thunk takes the original payload type (extracted using conditional infer types) as parameter; implementation either returns promise or not-promise, so the implementation will either pass thunk as resolve or call thunk as-is
\item server receive actions - thunk pattern as well
\item browser send action -- need a runtime flag to silently disable channel reuse (which would otherwise violate linearity) because the sending action is delayed by the Promise resolve
\item browser receive action -- thunk pattern
\end{itemize}

\begin{figure}[!h]
\begin{subfigure}{\textwidth}
\begin{lstlisting}[language=javascript,tabsize=2]
const [	label, payload, succ] = this.handler;
send(label, payload);
next(succ);
\end{lstlisting}
\caption{\dots}
\end{subfigure}
\hfill
\begin{subfigure}{\textwidth}
\begin{lstlisting}[language=javascript,tabsize=2]
const thunk = (label, payload, succ) => {
	send(label, payload);
	next(succ);	
};

if (this.handler instanceof Promise)
	this.handler.then(thunk);
else 
	thunk(this.handler);
\end{lstlisting}
\caption{\dots}
\end{subfigure}
\captionof{figure}{\dots}
\end{figure}

\begin{lstlisting}[language=javascript,tabsize=2]
(*@\hl{let used = false;}@*)
...
return class extends React.Component {
	render() {
		const props = {
			[eventLabel as string]: (event) => {
				(*@\hl{if (used) return;}@*)
				(*@\hl{used = true;}@*)

				const result = handler(event);
				if (result instanceof Promise) {
					result.then(send);				
				} else {
					send(result);				
				}
			}		
		}
	}
}
\end{lstlisting}

\subsection{Limitations}

Supporting asynchronous logic on top of our
existing implementation in \fancyname{ReactMPST} comes at a cost
of losing affine linearity guarantees.
If the send action handler is asynchronous, then the UI event handler
returns before the runtime transitions, so the UI event that can trigger
the send action is still active, which means the user can trigger
a channel linearity violation. This means we need a boolean flag
to keep track of channel usage for browser-side endpoints, 
as done in \cite{Hybrid2016}.

If we decouple the \texttt{sendMessage()} 
from the \texttt{advance()} function, we could advance to the successor state
immediately \textit{before} handling the asynchronous return value.
This restores affine channel linearity guarantees, but 
results in the UI being inconsistent with the channel
actions.