### 2020-06-11 report/sections/server/extensions/error.tex
Report: progress on chapter 6 error handling

\section{Error Handling}
\label{section:error}

A robust error handling framework is critical
in a distributed setting.
It is naive to assume that endpoint implementations
are entirely free from exceptions.
As Fowler pointed out in \cite{Exceptional},
session type implementations that do not account for failure
are of limited use in distributed programming.

We motivate the need for handling a variety exceptions in
modern web programming in \cref{subsection:errormotivation},
demonstrate how we implement a structured error handling
framework through extending the generated APIs and 
runtime in \cref{subsection:errorapi,subsection:errorruntime}
respectively, and explain the limitations of our framework
in \cref{subsection:errorlimit}.

\subsection{Motivation}
\label{subsection:errormotivation}

Using the \tprotocol{2FA} example from
\cref{lst:2FA}, the database lookup function
might rely on a cloud service, and will throw an exception
if the cloud service cannot be reached.
This may cause \trole{Svr} to crash.
From the perspective of session types, we need to
ensure that \trole{Client} isn't blindly
waiting for the session to continue if \trole{Svr}
threw an exception.

We classify this as an example where \trole{Svr}
terminates the session because of a \textit{logical error}
in its implementation.
Existing work \cite{Exceptional,AffineSessions} 
on error handling for session types address this
through extending the process calculus with an
\textit{explicit cancellation} operator to ensure that
channel resources are closed during exception handling.

However, the context of modern web programming
introduces new possibilities for session cancellation
that are not addressed by existing work.
Browser-side endpoints can disconnect from the session due
to network connectivity issues or simply by closing
the browser. Similarly, the server may also face connectivity
issues. In order to deliver a robust error handling
framework tailored for integrating session types in 
modern web programming, we must also take these variants
of session cancellation into account.

We also need a novel approach for defining cancellation handlers.
The work by Fowler et al. \cite{Exceptional} integrates 
exception handling for session types in a 
\textit{functional language}, where their design of
exception handlers build upon existing work on effect handlers.
Our work targets a non-functional language in TypeScript,
which means any piece of code can be ``effectful'',
so the exception handler API we generate for developers
need to respect this property.
Exception handlers should also be well parameterised
to empower the developer to handle cancellations more effectively
-- for instance, if the developer knows which endpoint
emitted the cancellation, appropriate clean-up operations
(such as reverting database changes or application-specific
logging) can be performed.

\subsection{API Extension}
\label{subsection:errorapi}

\begin{itemize}
\item developers just need to provide cancellation handlers along with the initial state
\item we standardise on parameterising it with role and reason: the reason is typed `any' to give developers the autonomy for implementation details
\end{itemize}

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
new ATMSvr(wss, logic, async (role: Roles.All, reason: any) => {
	if (role === Roles.Self) {
		// Handle internal server error	
	} else {
		// Handle client error:
		// revert incomplete transactions from database
		await db.revertTransactions(role);
	}
});
\end{lstlisting}
\captionof{lstlisting}{Example Cancellation Handler for Server Endpoint}
\label{lst:cancelsvr}
\end{figure}

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
<ATMClient
	endpoint='ws://localhost:8080'
	...
	cancellation={(role: Roles.All, reason: any) => (
		<div>
			{role === Roles.Self
				? <h1>Internal Error</h1>
				: <h1>Server Error</h1>}
			<p>{reason}</p>
		</div>
	)}	
/>
\end{lstlisting}
\captionof{lstlisting}{Example Cancellation Handler for Browser Endpoint}
\label{lst:cancelclient}
\end{figure}

\subsection{Runtime Extension}
\label{subsection:errorruntime}

The generated session runtime \dots

\begin{itemize}
\item build upon websocket RFC which specifies the `base' closing (or exit) code
\item we additionally define contracts for exit codes in our use case (e.g. role occupied, logical error), and make sure it does not override any websocket `traditions' as per the RFC
\item all based on well-defined event listeners on the websocket close and error events
\item must consider cancellation in two phases
\end{itemize}

\subsubsection{Cancellations during Session Joining}

\begin{itemize}
\item server -- emits role occupied
\item client -- closes browser whilst waiting for others
\end{itemize}

\subsubsection{Cancellations during EFSM Execution}

\begin{itemize}
\item server -- dies
\item server -- logical error
\item client -- dies
\item client -- logical error
\end{itemize}

\subsection{Limitations}
\label{subsection:errorlimit}

The expressiveness of our error handling framework is
limited by the expressiveness of the 
\lstonelinejs{try} / \lstonelinejs{catch} mechanism
in TypeScript.

\begin{itemize}
\item Cancellation handlers are not typed -- namely, we use `any';
\item exception handler blocks in typescript are not typed
\item the `logical error' idea is also very vague as all code can be effectful in typescript (statically); cannot specifically say that the implementation is expected to `send payload of type int, but possibly throw error of type Error', which would otherwise be possible in other languages(?)
\end{itemize}