### 2020-06-12 report/sections/evalconcl/eval/twobuyer.tex

\section{Routed Multiparty Sessions: \fancyname{Two Buyers}}
\label{section:evaltwobuyer}

Interested readers can find the full implementation on
GitHub\footnote{\url{https://github.com/ansonmiu0214/SessionTS-Examples/TwoBuyer}}.

\begin{itemize}
\item show that the routing aspect is transparent to both sides
\item same flexibilities apply -- we use react context for state management
\end{itemize}