### 2020-06-12 report/sections/appendix/eval.tex
Report: progress on chapter 10

\chapter{Artefacts for Evaluation}
\label{chapter:evalcode}

\section{Implementation for \tprotocol{Noughts and Crosses}}
\label{section:evalcodegame}

% BOARD CLASS

\section{Implementation for \tprotocol{Two Buyer}}
\label{section:evalcodetwobuyer}

\section{Implementation for \tprotocol{ATM}}
\label{section:evalcodeatm}