### 2020-06-12 report/sections/evalconcl/eval/twobuyer.tex
Report: complete chapter 10.2 sans listings

\section{Routed Multiparty Sessions: \tprotocol{Two Buyers}}
\label{section:evaltwobuyer}

We implement the \tprotocol{Two Buyer} protocol
introduced in \cref{section:twobuyer}. 
This protocol cannot be implemented using
existing proposals \cite{PureScript2019,MVU2020} for integrating
session types into web development.
\codegen overcomes the limitations of existing work
through implementing novel theory of routed multiparty
session types formalised in \cref{chap:theory}.

We present the implementation of the \trole{S}eller
in \cref{lst:evaltwobuyerseller}, 
and the implementation
of the peer-to-peer interaction by buyer \trole{A}
in \cref{lst:evaltwobuyerA}
The main point to note is that the routing mechanism
is completely transparent to the developer,
which shows the elegance of our solution.

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
// two buyer seller
\end{lstlisting}
\captionof{lstlisting}{\dots}
\label{lst:evaltwobuyerseller}
\end{figure}

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
// two buyer A p2p
\end{lstlisting}
\captionof{lstlisting}{\dots}
\label{lst:evaltwobuyerA}
\end{figure}

The \newtheory implementation does not affect
the compatibility of our generated APIs with external libraries.
The \trole{S}eller endpoint is set up as an Express.js
application, and both buyers still use the React Context API
for application state management. We show an example of
the latter in the developer's implementation for
buyer \trole{A} receiving and keeping track of the server's quote
using a React Context in \cref{lst:evaltwobuyercontext}.

\begin{figure}[!h]
\begin{lstlisting}[language=javascript]
// two buyer A context
\end{lstlisting}
\captionof{lstlisting}{\dots}
\label{lst:evaltwobuyercontext}
\end{figure}

We include the implementations of the remaining EFSM states
for buyer \trole{A} in \cref{section:evalcodetwobuyer}
for completeness;
interested readers can find the full implementations for the other
endpoints alongside the generated code on
GitHub\footnote{
\url{https://github.com/ansonmiu0214/SessionTS-Examples/TwoBuyer}
}.