### 2020-06-14 report/sections/general/theory/encoding.tex

\section{Router-Parameterised Encoding}
\label{section:encoding}

\begin{itemize}
\item define encoding on global types
\item define encoding on local types
\item define encoding on LTS actions
\item motivate with twobuyer example
\end{itemize}

\subsection{Correspondence between Global and Local Type Encodings}
\begin{itemize}
\item prove proposition
\item state corollary for subsequent proof
\end{itemize}

\subsection{Preserving Well-formedness}

\subsection{Preserving Communication}