### 2020-06-15 report/sections/background/interim/typescript.tex
Progress: start moving background

\section{TypeScript}

\subsection{Overview}
\begin{itemize}
\item Background
\item Aims
\item Example code snippet
\end{itemize}

\subsection{Gradual Type System}
\begin{itemize}
\item Introduce static (Java), dynamic (Python/JS)
\item Use pros and cons of the former two systems to motivate the concept of a gradual type system
\item Shwo example of unsoundness
\end{itemize}

\subsection{Language Features}
\begin{itemize}
\item Outline features that are either (a) relevant to implementing web applications, (b) differ from the features prevalent in traditional OO (Java) or functional (PureScript) languages
\end{itemize}

\subsubsection{Promises}
\textit{Relevant for modelling `state-specific input futures' from the hybrid Java paper?}

\subsubsection{Discriminated Union}
\textit{Useful for dealing with the switch case approach for branch receives?}

\subsubsection{Index Types}
\textit{With index types, you can get the compiler to check code that uses dynamic property names.}

\subsubsection{Conditional Types}
\textit{Compare this with the type dependencies achieved in the PureScript paper?}