### 2020-06-15 report/sections/background/related.tex

\section{Related Work}
\label{section:bgrelated}

\subsection{Scribble and Endpoint API Generation}
\label{subsection:bgscribble}

\subsection{Session Types in Web Development}
\label{subsection:sessiontypewebdev}