### 2020-06-16 report/sections/introduction/objectives.tex

\section{Objectives}

The objective of this project is to provide
developers with a development workflow
that provides communication safety guarantees in modern
web programming through multiparty session types.

We tackle the limitations of the current
state-of-the-art with a two-step strategy:

\subparagraph{S1}
\dots

\subparagraph{S2}
\dots

By step \textbf{S1}, we address challenge \textbf{C1}.

In step \textbf{S2}, we \dots

%an end-to-end framework for writing full-stack 
%applications that \textbf{statically}
%conform to a communication protocol, 
%
%With respect to the end goal of 
%offering developers a development workflow
%that provides communication safety guarantees in modern
%web programming through multiparty session types,
%we observe two main limitations from the state-of-the-art:

The objective of this project is to provide an end-to-end
framework for writing full-stack web applications that
\textbf{statically} conform to a communication protocol.
Developers specify their communication protocol using
the Scribble protocol description language \cite{Scribble}
which is based on multiparty session type (MPST) theory \cite{MPST}.
The Scribble protocol is used as a basis for generating
TypeScript APIs targeted for the Node.js server-side runtime
and React.js browser-side framework. 
By implementing the generated APIs in their full-stack applications,
developers can enjoy \textbf{static} guarantees with respect 
to communication safety: there will not be any communication mismatches,
and channel linearity will be respected.
Finally, we extend our framework to support communication
protocols that cannot be natively implemented over
a server-centric network topology using WebSockets.
This involves defining a theory of routed multiparty session types
to formalise the preservation of communication safety properties.

%Among the many languages that compile to JavaScript,
%TypeScript stands out as the most intuitive to use as it is defined
%to be a \textit{superset} of JavaScript. It provides developers
%with type-safety through its gradual, structural type system.
%Whilst some \cite{MVU2020} point out this limits its usability for
%encoding multiparty session types, we believe that the language
%offers sufficient features that we can use to provide developers with
%communication safety guarantees whilst preserving a flexible, natural
%and idiomatic workflow. By building our work upon TypeScript, we
%work towards incorporating MPST into mainstream web development,
%which reduces development time by programmatically verifying 
%implementations for communication correctness.