### 2020-06-17 report/sections/server/codegen.tex
Report: review chapter 3

\chapter{\codegen: Session Type API Generation for TypeScript}
\label{chap:codegen}

In this chapter, we present
\codegen,
a toolchain for generating TypeScript APIs that developers can
use to write web applications that conform to their specified
communication protocol.
The toolchain is publicly available at \cite{repo}.

\codegen supports communication protocols that define 
\emph{server-centric} topologies, meaning that: 
\textbf{(1)} there is exactly one participant executed on the 
Node.js runtime; 
\textbf{(2)} all other participants run on the web browser; and
\textbf{(3)} all non-server participants only communicate with the server.
An example would be the \tprotocol{Noughts and Crosses}
multiplayer game (\cref{lst:game}).
We relax this assumption in \cref{part:general}.

\input{sections/server/codegen/workflow}

\input{sections/server/codegen/implementation}

\input{sections/server/codegen/testing}