### 2020-06-17 report/sections/server/node.tex
Report: review chapter 4

\chapter{\nodecodegen: Back-End Session Type Web Development}
\label{chap:node}

In this chapter we present \nodecodegen,
our session type API generation strategy for 
server-side endpoints implemented on the Node.js runtime 
\cite{nodejs}.
We continue to use the running example of the \tprotocol{Adder}
protocol (\cref{lst:adder}),
and refer to the FSM of the \trole{Svr} endpoint
(\cref{fig:addersvrfsm}) throughout this chapter.

We discuss the challenges of implementing
session types on the Node.js runtime 
(\cref{section:nodechallenges})
and motivate our approach (\cref{section:nodeapproach}).
We explain our design choices for encoding
the EFSM for server-side endpoints (\cref{section:nodeefsm}),
and detail the session runtime that is designed
to execute the EFSM on Node.js (\cref{section:noderuntime}).
We conclude by analysing some alternative designs 
(\cref{section:nodealt}) and evaluating the limitations 
of our approach (\cref{section:nodelimitations}).

\begin{figure}[!b]
\centering
\includegraphics[width=0.6\textwidth]{AdderSvrFSM}
\captionof{figure}{\trole{Svr} Endpoint FSM
in \tprotocol{Adder} protocol}
\label{fig:addersvrfsm}
\end{figure}

\input{sections/server/node/challenges}
\input{sections/server/node/approach}
\input{sections/server/node/efsm}
\input{sections/server/node/runtime}
\input{sections/server/node/alt}
\input{sections/server/node/limitations}

\section{Summary}
In this chapter, we have presented
our session type API generation strategy
which targets the Node.js runtime for server-side
endpoints.

We motivated our approach of generating
handler-style APIs that act as seams in the session runtime
which executes the EFSM.
By doing so, the mechanism for sending and receiving
messages are not exposed to the developer,
which makes channel reuse impossible by construction.

We discussed our session runtime implementation in great
depth and iterated upon the design to arrive at 
a minimal runtime implementation that executes the EFSM
using a message-passing abstraction internally as well.
In particular, we motivated how to handle message receives
in a way that preserves the message ordering specified
in the communication protocol.

Developers that implement their Node.js endpoint application
using the generated APIs enjoy protocol conformance
by construction: 
the TypeScript type system will prevent type mismatches
in the handlers defined by the developer
(e.g. sending a \lstonelinejs{string} instead of 
a \lstonelinejs{number}, or using a message label
that is not defined on the EFSM state),
and the handler-style APIs themselves conceal channel
resources to prevent channel reuse.