### 2020-06-17 report/sections/server/node/challenges.tex

\section{Challenges}
\label{section:nodechallenges}

Our goal with \nodecodegen is to
generate session type implementations that
provide as much \textbf{static} communication safety guarantees
as possible,
in a manner suitable for both the TypeScript language 
and the single-threaded event loop model in Node.js
TypeScript's type system
is neither affine nor linear, so we need to work around the 
language's limitations to enforce linear channel usage.

We also aim to provide the developer with
idiomatic event-driven APIs commonly found in
server-side TypeScript development.