### 2020-06-17 report/sections/appendix/eval.tex
Report: feedback

\chapter{Artefacts for Evaluation}
\label{chapter:evalcode}

\section{Implementation for \tprotocol{Noughts and Crosses}}
\label{section:evalcodegame}

\begin{lstlisting}[language=javascript,tabsize=2]
/**
 * User implementation of game board state.
 * Export game board instance.
 */

import { Coordinate as Point } from './Game/GameTypes';

enum Cell { Empty, P1, P2 };
export enum MoveResult { Win, Draw, Continue };

class Board {

	private _board: Array<Array<Cell>>;
	private _emptyCellCount: number;

	constructor() {
		this._board = [[Cell.Empty, Cell.Empty, Cell.Empty],
		[Cell.Empty, Cell.Empty, Cell.Empty],
		[Cell.Empty, Cell.Empty, Cell.Empty]];
		this._emptyCellCount = 9;
	}

	// Factory for generating a function that, given a player role,
	// places the marker and returns the game result with respect to 
	// that player.
	private _makeMove = (marker: Cell) => ({ x: row, y: col }: Point) 
		=> {
		return new Promise<MoveResult>((resolve, _) => {
			setTimeout(() => {
				// Update board state
				this._board[row][col] = marker;
				this._emptyCellCount--;

				// Check for winning move
				if (this._board[row].every(cell => cell === marker))
					resolve(MoveResult.Win);

				/* Winning column */
				if (this._board.every(row => row[col] === marker))
					resolve(MoveResult.Win);

				/* Placed middle marker - check corners */
				if (row === 1 && col === 1) 
					if ((this._board[0][0] === marker 
						&& this._board[2][2] === marker) ||
						(this._board[0][2] === marker 
						&& this._board[2][0] === marker))
						resolve(MoveResult.Win);

				/* Placed corner marker - check diagonals */
				if (row !== 1 && col !== 1) 
					if (this._board[1][1] === marker 
						&& this._board[2 - row][2 - col] === marker)
						resolve(MoveResult.Win);

				resolve(this._emptyCellCount === 0 ? 
					MoveResult.Draw : MoveResult.Continue);
			}, 2000);
		})

	}

	p1(move: Point) { return this._makeMove(Cell.P1)(move); }
	p2(move: Point) { return this._makeMove(Cell.P2)(move); }

	clear() {
		this._board = [[Cell.Empty, Cell.Empty, Cell.Empty],
		[Cell.Empty, Cell.Empty, Cell.Empty],
		[Cell.Empty, Cell.Empty, Cell.Empty]];
		this._emptyCellCount = 9;
	}

}

// Initialise state
export const board = new Board();
\end{lstlisting}

% BOARD CLASS

%\section{Implementation for \tprotocol{Two Buyer}}
%\label{section:evalcodetwobuyer}

\section{Package Dependencies for Benchmarks}
\label{section:evalbenchmarkpackagejson}

\begin{lstlisting}[language=javascript,tabsize=2,
title=Packages for Ping Pong Client]
{
  "dependencies": {
    "@testing-library/jest-dom": "^4.2.4",
    "@testing-library/react": "^9.3.2",
    "@testing-library/user-event": "^7.1.2",
    "@types/jest": "^24.0.0",
    "@types/node": "^12.0.0",
    "@types/react": "^16.9.0",
    "@types/react-dom": "^16.9.0",
    "react": "^16.13.1",
    "react-dom": "^16.13.1",
    "react-scripts": "3.4.1",
    "typescript": "~3.7.2"
  },
  "devDependencies": {
    "@babel/preset-typescript": "^7.10.1",
    "customize-cra": "^1.0.0",
    "react-app-rewired": "^2.1.6"
  }
}
\end{lstlisting}

\begin{lstlisting}[language=javascript,tabsize=2,
title=Packages for Ping Pong Server]
{
  "dependencies": {
    "express": "^4.17.1",
    "ws": "^7.3.0",
    "zombie": "^6.1.4"
  },
  "devDependencies": {
    "@types/express": "^4.17.6",
    "@types/ws": "^7.2.4",
    "ts-node": "^8.10.2",
    "typescript": "^3.9.3"
  }
}
\end{lstlisting}