### 2020-06-17 report/sections/server/react/challenges.tex

\section{Challenges}
\label{section:reactchallenges}

Our goal with \fancyname{ReactMPST} is to
generate session type implementations for the web browser.
Integrating session types into user interfaces is inherently difficult --
assuming channel actions are bound to user interface events 
(i.e. clicking a button sends a message), how does one formalise
channel linearity? How do we guarantee that, if the button
triggers a send at some EFSM state, that it triggers not more than
one channel action, \textit{and} the user cannot trigger the action
at another EFSM state?

We recap how existing work 
(as discussed in \cref{subsection:sessiontypewebdev})
tackle browser-side session typing.
Fowler introduced the concept of model types to prevent
channel linearity violation in his proposal for
integrating session types with
graphical user interface (GUI) programming \cite{MVU2020}, but
it uses the Links web programming language \cite{LINKS} which
lacks compatibility with the ecosystem of JavaScript
libraries that one might use in front-end development as well.
The session type-safe web development framework presented
by King et al. \cite{PureScript2019} generates APIs for
a functional target language in PureScript, and relies on
the \textit{Concur UI} framework that constructs UIs sequentially.

As motivated in \cref{section:intro},
we find these proposals to come at the cost of
limiting developer productivity by adopting unconventional practices
that may require a learning curve.
Through our work, we aim to distil the key
concepts from existing work that provide session type safety
for web-based GUI programming, and implement them using
mainstream front-end web development tools (i.e. TypeScript
and React), to provide developers with an intuitive way to implement
browser-side endpoints that guarantee communication safety.