### 2020-06-17 report/sections/server/react/limitations.tex

\section{Limitations}
\label{section:reactlimitations}

Our EFSM encoding provides \textbf{affine} channel
usage guarantees for send states as supposed to \textit{linear} usage
guarantees. Suppose we bind a send action to a
button click event -- strictly speaking,
even if we guarantee that the button can be clicked exactly once,
we cannot guarantee that the button will ever be clicked by the user.
However, existing work \cite{PureScript2019,MVU2020} 
share the same limitation, and we view this as an
inherent limitation in front-end development, since there is
a limit to the types of assumptions we can make about how the
user interacts with the browser.
Redefining the definition of channel linearity 
for GUI programming use cases
would be an interesting discussion.

Our runtime implementation also limits the developer from
passing data into their EFSM state component through props.
This is because the runtime takes full control of which
state component to render, and by definition, what props 
are passed to the EFSM state components.
However, we equally view this as a way to encourage developers
to decouple communication state management from
business logic state management, in the spirit
of separating concerns.
In \cref{chap:eval},
we present examples of developer implementations
that use separate abstractions (e.g. the \textit{React Context API}) 
for propagating 
stateful application data in \cref{chap:eval}.