### 2020-06-18 report/sections/evalconcl/eval/benchmarks.tex

\section{Performance Benchmarks}
\label{section:benchmarks}

Whilst web applications implementing 
our generated APIs enjoy
communication safety guarantees, 
the presence of the session runtime acts
as an additional layer of abstraction between 
the application logic and the
WebSocket transport, which presents
a performance trade-off; we observe an
increase in the time taken to perform
each channel action, as a result of
the session runtime intercepting channel actions
and performing additional logic to ensure
that the endpoint conforms to the protocol.

\begin{figure}[!ht]
\begin{lstlisting}[language=Scribble]
global protocol PingPong(role Client, role Svr) {
	PING(number) from Client to Svr;
	choice at Svr {
		PONG(number) from Svr to Client;
		do PingPong(Client, Svr);
	} or {
		BYE(number) from Svr to Client;	
	}
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Ping Pong} Protocol}
\label{lst:pingpong}
\end{figure}

To measure the overhead of our implementation, we compare the
execution time of web-based implementations of the
Ping Pong protocol (\cref{lst:pingpong}) with and without
our generated APIs.

We parameterise the \tprotocol{Ping Pong} protocol by $n > 0$, the number of
round-trip messages. This is standardised in the application logic
across experiments.
Upon establishing a connection, the experiment proceeds as follow:

\begin{enumerate}

\item \trole{Client} sends \tmsg{PING($m$:number)} to \trole{Svr}, 
with $m = 0$ initially.

\item \trole{Svr} receives \tmsg{PING($m$:number)}, and
conditionally responds based on $n$:

\begin{enumerate}
\item If $m + 1 < n$, then \trole{Svr} replies \tmsg{PONG($m + 1$)}.
\trole{Client} responds to \tmsg{PONG} by returning to
step 1 with $m$ set as the payload from \tmsg{PONG}.

\item Otherwise, $m + 1 = n$, then \trole{Svr} responds with 
\tmsg{BYE($m + 1$)}, as $n$ round trips have taken place. 
\trole{Client} responds to \tmsg{BYE} by 
closing the connection, thus ending the experiment.
\end{enumerate}

\end{enumerate}

We note that the \tprotocol{Ping Pong} protocol 
implements a \textit{binary} session. 
It would be interesting to observe the overhead in a \textit{multiparty}
context, but due to limited time constraints,
our benchmarking suite does not support multiple browser targets.
Benchmarking multiparty protocols would also require writing multiple
distinct React applications using the generated APIs -- as this is currently
a manual process, doing this for multiple roles requires more time than
available.

\subsection{Setup}

In order to measure the overhead as accurately as possible,
we outline the logic that all implementations must follow:

\begin{figure}[!ht]
\centering
\includegraphics[width=0.5\textwidth]{pingpongclient}
\captionof{figure}{User Interface of \trole{Client} Endpoint in
\tprotocol{Ping Pong} Protocol}
\label{fig:pingpongclient}
\end{figure}

\subparagraph{\tprotocol{Ping Pong} \trole{Client} on React:}
\begin{itemize}

\item All \trole{Client}s implement the same user interface 
(\cref{fig:pingpongclient}), rendering
a \texttt{<button>} which triggers the send, and
a \texttt{<div>} captioned with the number of \tmsg{PONG}s received.

\item \trole{Client}s will use the React Context API \cite{reactcontext}
for application state management, i.e. the number of \tmsg{PONG}s received. 
We wrap the session logic in a \texttt{<Benchmark>} component 
that acts as the \texttt{ContextProvider} using its component state.

\item To automate the benchmark, we use the React Refs API \cite{reactrefs}
to access the DOM \texttt{<button>} node programmatically, in order to
simulate the click event and send a \tmsg{PING} message upon establishing
the WebSocket connection, or upon receiving a \tmsg{PONG}.

\item We use the production build generated by 
\textit{create-react-app} \cite{cra} for all experiments, which performs the
compilation into JavaScript. We serve the production build using the
\textit{serve} package \cite{npmserve} available on \textit{npm}.

\end{itemize}

\subparagraph{\tprotocol{Ping Pong} \trole{Svr} on Node:}
\begin{itemize}

\item We use the built-in \texttt{console.time} function to record
the execution time of all experiments. 
The timer starts when a WebSocket connection has been established at
\trole{Server}, and stops when on a \texttt{CloseEvent}.

\item To observe the execution pattern, the \trole{Svr} will log the running 
elapsed time for every \tmsg{PING} message received.

\item All \trole{Svr}s run the benchmarks without a real web browser,
using headless browsing functionality from 
the \textit{Zombie.js} \cite{zombie} package.

\item \trole{Svr} logic is parameterised by the number of round trips,
$n$, configured through an environment variable passed through 
the command line.

\item We use the compiled JavaScript versions of all \trole{Svr}s
for the experiments.

\end{itemize}

Code is available at
\url{https://github.com/ansonmiu0214/SessionTS-Benchmarks}.
Interested readers may follow the \texttt{README.md}
to run the benchmarks and visualise the logs using the 
interactive notebook in the same directory.

We run the experiments under a network of latency 0.165ms
(64 bytes ping), and repeat each experiment 20 times.
Execution time measurements  are taken using a machine 
equipped with Intel i7-4850HQ CPU (2.3 GHz, 4 cores, 8 threads), 
16 GB RAM, macOS operating system version 10.15.4, 
Node.js runtime version 12.12.0, and
TypeScript compiler version 3.7.4.
We standardise all packages used in the front- and back-end
implementations across experiments. Details can be found in
\cref{section:evalbenchmarkpackagejson}.

The benchmark compares three implementations of 
the \tprotocol{Ping Pong} protocol:

\subparagraph{bare:}
The \texttt{bare} implementation directly interfaces with 
WebSocket primitives for sending and receiving. 
The implementation executes the \tprotocol{Ping Pong} protocol, but does not 
guarantee communication safety by construction -- e.g. the user can click
the \tmsg{PING} button multiple times before a \tmsg{PONG} message 
is received, violating channel linearity. 
This represents the typical developer implementation without using the
MPST framework.

\subparagraph{bare_safe:}
The \texttt{bare_safe} implementation also directly interfaces with
WebSocket primitives for communication, but assumes the developer
implements minimal viable workarounds to address the lack of
communication safety. Here, the developer renders an inactive version
of the \tmsg{PING} button when the \tmsg{PING} message has been sent
but a response has yet to be received; a \texttt{visible} boolean flag
is used to explicitly manage which \texttt{<button>} to render.

\subparagraph{mpst:}
The \texttt{mpst} implementation uses the APIs generated from 
\fancyname{SessionTS}, so it enjoys the
communication safety guarantees from our methodology.

\subsection{Execution Pattern}
\label{section:execpattern}

We compare the execution patterns of exchanging 10,000 Ping-Pongs
throughout 20 repeated experiments across the three implementations.
We visualise the elapsed time with respect to the number of \tmsg{PING}s
received in \cref{fig:execution}: each line represents one
execution of the benchmark.

\begin{figure}[!ht]
\centering
\begin{subfigure}[b]{0.49\textwidth}
\centering
\includegraphics[width=\textwidth]{execbare10000}
\caption{\texttt{bare}}
\label{fig:executionbare}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.49\textwidth}
\centering
\includegraphics[width=\textwidth]{execbaresafe10000}
\caption{\texttt{bare_safe}}
\label{fig:executionbaresafe}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.49\textwidth}
\centering
\includegraphics[width=\textwidth]{execmpst10000}
\caption{\texttt{mpst}}
\label{fig:executionmpst}
\end{subfigure}
\caption{Comparison of Execution Pattern for 10,000 Ping-Pongs}
\label{fig:execution}
\end{figure}

Relative to the two \texttt{bare} implementations, the
\texttt{mpst} version performs more consistently, perhaps as a result
of the session runtimes handling all WebSocket interactions in a
systematic way. The gradient of the graph represents the rate at which
a Ping-Pong round trip takes place. 
We observe a steeper gradient when the protocol begins, 
which illustrates the overhead incurred in the 
session joining phase in our generated APIs.
Aside from these factors, all three implementations generally share
similar characteristics in their protocol execution.

\subsection{Overhead}
The execution time is the total time taken
for the \trole{Client} to connect to the \trole{Server}
\textit{and} complete the parameterised number of round trips.
We compare the total execution time (\textit{Exec. Time}) 
and execution time per round trip (\textit{Exec. Time / Ping-Pong}) -- 
averaged over 20 repeated experiments -- across the three implementations,
for $n \in \{10^2, 10^3, 10^4\}$.
We summarise the results in \cref{table:overhead}.

\renewcommand{\arraystretch}{1.5}
\begin{table}[!ht]
\centering
\begin{tabular}{||c||c|c|c||c|c|c||}
\hline
\multirow{2}{*}{$n$} & 
\multicolumn{3}{c||}{Exec. Time} & 
\multicolumn{3}{c||}{Exec. Time / Ping-Pong} \\
\cline{2-7}
 & \texttt{bare} & \texttt{bare_safe} & \texttt{mpst} 
 & \texttt{bare} & \texttt{bare_safe} & \texttt{mpst} \\
\hline\hline
$10^2$ & 89.64ms & 107.09ms & 186.23ms & 0.90ms & 1.07ms & 1.86ms \\
$10^3$ & 642.92ms & 663.91ms & 1155.48ms & 0.64ms & 0.66ms & 1.16ms \\
$10^4$ & 3542.16ms & 3837.97ms & 7015.25ms & 0.35ms & 0.38ms & 0.70ms \\
\hline
\end{tabular}
\captionof{table}{Comparison of Execution Time for 
100, 1,000 and 10,000 Ping-Pongs}
\label{table:overhead}
\end{table}
\renewcommand{\arraystretch}{1}

We note that the addition of a session runtime for all roles in the 
\texttt{mpst} implementation \textit{does} incur a performance overhead. 
This is made apparent when looking closely at 
\textit{Exe. Time / Ping-Pong};
we visualise this in \cref{fig:timepermsg}.

\begin{figure}[!ht]
\centering
\includegraphics[width=\textwidth]{timepermsg}
\captionof{figure}{Comparing Average Time per Ping-Pong 
Across Implementations}
\label{fig:timepermsg}
\end{figure}

The \texttt{mpst} implementation records greater round trip times compared
to both \texttt{bare} and \texttt{bare_safe} variants. This is expected,
as the runtime for \textit{each} role performs additional logic during
\textit{both} the sending and receiving of messages. 
For example, receiving a message involves performing
(albeit $O(1)$ time complexity) operations on the message queue and
handler queue. As for browser-side implementations, every EFSM transition
invokes an updated \texttt{render()} on the VDOM, which requires
reconciliation internally by React to update the browser DOM accordingly.

We also observe that the (round trip time) performance gap between 
the \texttt{mpst} implementation and the two 
\texttt{bare} implementations narrows as the number of round trips 
increase. This suggests that the session joining logic in our
implementation yields greater overhead than the additional processing
logic injected during communication, which is consistent with 
our findings in \cref{section:execpattern}.

We interpret the overhead as a trade-off between maximising performance and
maximising the \textbf{static} communication safety guarantees for
web applications. In particular, the event-driven nature of 
browser-side logic makes it highly challenging to guarantee 
communication safety, as having ``active'' channel actions on the DOM
allows linearity to be violated by the user, even if one verifies
linear channel usage in the source code. The \texttt{mpst} implementation
properly addresses the problem by statically guaranteeing,
by construction, that the UI component rendered on the DOM will only
contain channel actions permitted at that state in the EFSM execution.

One may argue that the \texttt{bare_safe} implementation also
provides said guarantee but adds negligible overhead. 
By inspecting the workaround in the \texttt{bare_safe} implementation
(\cref{lst:workaround}),
it is clear that this does not scale for more complicated protocols with
larger EFSMs.

\begin{figure}[!ht]
\begin{lstlisting}[language=javascript, tabsize=2]
click() {
	this.state.ws?.send(JSON.stringify({
		label: 'PING',
		payload: [this.context.count],
	}));
	this.setState({ visible: false });
}

render() {
	return this.state.visible 
		? <button
		   ref={this.state.button}
		   onClick={this.click.bind(this)}
		   >Ping</button>
		: <button>Ping</button>;
}
\end{lstlisting}
\captionof{lstlisting}{Preventing Channel Linearity Violation in
\texttt{bare_safe} \tprotocol{Ping Pong}}
\label{lst:workaround}
\end{figure}

The workaround is tailored to the \tprotocol{Ping Pong} protocol
and toggles the \texttt{visible} flag to hide the button that triggers
a send action when \trole{Client} transitions to a receive state.
In fact, this ``workaround'' precisely generalises to our 
\texttt{mpst} implementation of having some form of wrapper
component that renders the UI component corresponding to the
current EFSM state. Our API generation strategy formally implements
the EFSM for browser-side logic.

We also compare the server-side logic between the \texttt{base} and
\texttt{mpst} implementations in \cref{fig:nodepingpong}, and observe
that our generated APIs (\cref{lst:nodepingpongmpst}) 
allow the developer to focus on the
implementation detail, whilst the naive bare implementation 
(\cref{lst:nodepingpongbare}) chooses to interleave communication logic
with application logic, which arguably contributes towards a hidden
source of bugs found when implementing more complex protocols.

\begin{figure}[!ht]
\centering
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
socket.onmessage = ({ data }) => {
	const { label, payload } = JSON.parse(data.toString());
	if (label === 'PING') {
		let count: number = payload[0];
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			socket.send(JSON.stringify({
				label: 'BYE',
				payload: [count],
			}));
		} else {
			socket.send(JSON.stringify({
				label: 'PONG',
				payload: [count],
			}));
		}
	} else { throw new Error(`Unrecognised label: ${label}`); }
}
\end{lstlisting}
\caption{\texttt{bare} Implementation of \tprotocol{Ping Pong} \trole{Svr}}
\label{lst:nodepingpongbare}
\end{subfigure}
\hfill
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
const logic = new Implementation.S14({
	PING: (count) => {
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			return new Implementation.S16([
				Labels.S16.BYE, [count], new Implementation.S15()
			]);
		} else {
			return new Implementation.S16([
				Labels.S16.PONG, [count], logic
			]);
		}
 	}
});
\end{lstlisting}
\caption{\texttt{mpst} Implementation of \tprotocol{Ping Pong} \trole{Svr}}
\label{lst:nodepingpongmpst}
\end{subfigure}
\caption{Comparison of \texttt{bare} and \texttt{mpst}
Implementations for \tprotocol{Ping Pong} \trole{Svr}}
\label{fig:nodepingpong}
\end{figure}