### 2020-06-18 report/sections/server/react.tex

\chapter{\reactcodegen: Front-End Session Type Web Development}
\label{chap:react}

In this chapter, we present \reactcodegen,
our session type API generation strategy for browser-side endpoints
implemented using React \cite{React}.
We will refer to the FSM of the \trole{Client}
endpoint (\cref{fig:adderclientfsm})
from the \tprotocol{Adder} protocol 
throughout this chapter.

We highlight the challenges of session-typed GUI programming
(\cref{section:reactchallenges}) and motivate 
our approach (\cref{section:reactapproach})
for adapting the proposal from
\cite{MVU2020} for the React.js framework, along with a brief
introduction to the framework.
We proceed to explain how we leverage React
to encode the EFSM (\cref{section:reactefsm})
and implement the session runtime (\cref{section:reactruntime}),
focusing on our novel approach for preventing
channel reuse in GUI programming using React.
We conclude by analysing alternative designs
(\cref{section:reactalt})
and evaluating the limitations of our approach
(\cref{section:reactlimitations}).

\begin{figure}[!b]
\centering
\includegraphics[width=0.6\textwidth]{AdderClientFSM}
\captionof{figure}{\trole{Client} Endpoint FSM
in \tprotocol{Adder} Protocol}
\label{fig:adderclientfsm}
\end{figure}

\input{sections/server/react/challenges}
\input{sections/server/react/approach}
\input{sections/server/react/efsm}
\input{sections/server/react/runtime}
\input{sections/server/react/alt}
\input{sections/server/react/limitations}

\section{Summary}
In this chapter, we have presented
our session type API generation strategy
which targets the React framework for browser-side endpoints.

We motivated our approach of
implementing the concept of model types introduced in \cite{MVU2020}
using a popular front-end framework in React.
We define an interpretation of model types
using the EFSM obtained from a Scribble protocol.

We discussed how to prevent channel reuse in
GUI programming in React. We build upon the design choices
implemented in \nodecodegen by concealing channel resources
in the session runtime and exposing seams for the developer
to customise what interface to render at each EFSM state,
what user interface events are allowed to trigger a send action,
and how to handle messages received from the channel.

% Developers that implement...