### 2020-06-18 report/sections/acknowledgements.tex
Report: ack

\begin{abstract}

I would like to thank Prof. Nobuko Yoshida,
Fangyi Zhou and Dr. Francisco Ferreira for their
continuous support, encouragement and motivation
during the project. 
I extend my gratitude to the members
of the Mobility Reading Group.
Their regular meetings have been a source
of inspiration for me, and helped me spark new ideas
to incorporate into the project.
I would also like to thank Prof. Peter Pietzuch
for supporting me as my personal tutor
throughout the degree.
\\

I would like to thank all the friends I have made
throughout my time at Imperial.
In particular, to the friends under the namespace
labelled \textit{Reply 404} at the time of writing:
I am forever grateful for every bit of shared memory
over the past four years.
\\

Finally, I would like to thank my family
for their unconditional love and support throughout my life,
and encouraging me to pursue a degree abroad.

\end{abstract}